---
title: "Mini Ocean Lighting Project Stats: Confounding Factors"
collaborators: Savanna Saunders, Renee Grambihler
project: Mini Ocean Light Data
institution: Biosphere 2 Ocean
description: "Uses a weighted least squares multiple linear regression model to analyze how size, weight, and color changes in corals grown in Mini Ocean under different lighting treatments from the Summer 2024 to Spring 2025 Coral Lighting Experiment were affected by environmental factors"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

**Input**: 
  `CoralColorWeightData_Rates.csv`: spreadsheet of coral data with Before and After values subtracted from each other and then divided either by 3 or 6 depending on what Trial the data is from in order to get a weekly rate of change.
  
  `All-Environmental-Variables-Wide.xlsx`: spreadsheet of all environmental data (nutrients, water quality, DLI) from MO and R2 during all trials in wide format
  
  `All-Environmental-Variables-Long.xlsx`: spreadsheet of all environmental data (nutrients, water quality, DLI) from MO and R2 during all trials in long format

**Output**: 
  `ConfoundingFactorTesting-KruskalWallis-InitialResults.xlsx`: spreadsheet of significant Kruskal Wallis results testing if environmental variables differ between trials 
  
  `ConfoundingFactorTesting-KruskalWallis-EffectSize.xlsx`: spreadsheet of the effect size of the significant Kruskal Wallis results 
  
  `ConfoundingFactorTesting-Dunn.xlsx`: spreadsheet of Dunn Test results showing comparisons between all trials for environmental variables 
  
  `ConfoundingFactorTesting-LinearRegressions.xlsx`: spreadsheet of regular linear regression results testing if the environmental variables affected the trial outcomes
  
  `ConfoundingFactorTesting-WeightedLeastSquareRegressions.xlsx`: spreadsheet of weighted least squares linear regression results testing if the environmental variables affected the trial outcomes 


## Import Libraries
```{r import-libraries}
library(tidyverse)
library(readxl)
library(writexl)
library(rstatix)
library(lubridate)
library(lmtest)
library(modelr)
library(ggpubr)
# get rid of scientific notation
options(scipen = 9999)
# avoid dplyr error message
options(dplyr.summarise.inform = FALSE)  
```

# Import Data
Import Coral Data and Environmental Data: nutrients, water quality, DLI. 
```{r import-data}
# coral data
# rate of change per week
dfCoral_rates <- read_csv("InputData/Clean/CoralData/CoralColorWeightData_Rates.csv") %>%
  # take out PBR 4 because it isn't in the environemntal data files
  filter(Trial != "PBR 4" & Trial != "PBR 4 Control") 


# environmental data
# wide version
dfAllEnvVar <- read_excel("InputData/Clean/EnvironmentalData/All-Environmental-Variables-Wide.xlsx") %>%
  #remove nitrate HR since nitrate MR is scaled more for smaller values
  select(-`Nitrate HR`) %>%
  # remove weird temp outlier
  filter(`Temperature (ºC)` > 0)

# long version
dfAllEnvVar_long <- read_excel("InputData/Clean/EnvironmentalData/All-Environmental-Variables-Long.xlsx") %>%
  #remove nitrate HR since nitrate MR is scaled more for smaller values
  filter(variable_names != "Nitrate HR") %>%
  # remove weird temp outlier 
  mutate(variable_values = ifelse(variable_names == "Temperature (ºC)" & variable_values < 0, NA, variable_values))

```

# Preliminary Plotting
Just want to get an idea of the general layout of the data by Trial.  
## Coral 
```{r data-wrangling-coral-boxplots}
# boxplots
var_names_coral <- colnames(dfCoral_rates)[5:7]
plotList_coral <- list()

for (i in 1:length(var_names_coral)){
  plotList_coral[[i]] <- dfCoral_rates %>%
    mutate(Trial = factor(Trial, levels = c("PBR 1", "PBC 1", "PBR 2", "PBR 3", "PBC 2", "PBR 4", "PBC 3", "Control"))) %>%
    pivot_longer(cols = 5:7,
                 names_to = "variable_name",
                 values_to = "variable_value") %>%
    filter(variable_name == var_names_coral[i]) %>%
    ggplot(aes(x = Trial,
               y = variable_value,
               fill = Treatment)) +
    geom_boxplot() +
    ylab(var_names_coral[i]) +
    theme_bw() + 
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          panel.background = element_blank())+
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.margin = margin(0.25, 0.45, 0.25, 0.25, "inch"))+
    theme(strip.background = element_rect(fill = "white"))
  
  print(plotList_coral[[i]])
}

```

## Environmental Variables
```{r data-wrangling-env-vars-boxplots}
# boxplots
var_names_env <- unique(dfAllEnvVar_long$variable_names)
plotList_env_var <- list()

for (i in 1:length(var_names_env)){
  plotList_env_var[[i]] <- dfAllEnvVar_long %>%
    filter(variable_names == var_names_env[i]) %>%
    mutate(Trial = factor(Trial, levels = c("PBR 1", "PBC 1", "PBR 2", "PBR 3", "PBC 2", "PBR 4", "PBC 3", "Control"))) %>%
    ggplot(aes(x = Trial,
               y = variable_values,
               fill = Treatment)) +
    geom_boxplot() +
    ylab(var_names_env[i]) +
    # scale_x_date(date_breaks = "1 month",
    #              date_labels = "%b %Y") +
    theme_bw() + 
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          panel.background = element_blank())+
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.margin = margin(0.25, 0.45, 0.25, 0.25, "inch"))+
    theme(strip.background = element_rect(fill = "white"))
  
  print(plotList_env_var[[i]])
}

```

# Confounding Factor Testing
## Do the environmental factors vary significantly between trials? 
### Kruskal-Wallis
```{r env-fact-vary-trial-kruskal-experimental}
# kruskal-wallis test to see which variables are significantly different by trial
kruskal_all <- dfAllEnvVar_long %>%
  group_by(variable_names) %>%
  kruskal_test(variable_values ~ Trial)%>%
  filter(p < 0.05) %>%
  mutate(p = round(p, digits = 4))

# kruskal-wallis effect size to see how large the effect size is for each variable 
kruskal_all_effsize <- dfAllEnvVar_long %>%
  group_by(variable_names) %>%
  kruskal_effsize(variable_values ~ Trial) %>%
  mutate(effsize = round(effsize, 2))

```

Significant p-value from Kruskal-Wallis test: 
- Alkalinity
- Ammonia
- Calcium
- Chlorophyll 
- DLI 
- ORP
- pH
- Phycoerytherin
- Salinity 
- Temperature

Effect size of significant variables based on Kruskal-Wallis test: 
- Alkalinity: 0.05 = Small
- Ammonia: 0.02 = Small
- Calcium: 0.06 = Moderate
- Chlorophyll: 0.20 = **Large**
- DLI: 0.20 = **Large**
- ORP: 0.15 = **Large**
- pH: 0.59 = **Large**
- Phycoerytherin: 0.29 = **Large**
- Salinity: 0.42 = **Large**
- Temperature: 0.39 = **Large**

4% of variance in Alkalinity is explained by Trial. 
2% of variance in Ammonia is explained by Trial. 
6% of variance in Calcium is explained by Trial.
20% of variance in Chlorophyll is explained by Trial.
20% of variance in DLI is explained by Trial.
15% of variance in ORP is explained by Trial.
59% of variance in pH is explained by Trial.
29% of variance in Phycoerytherin is explained by Trial.
42% of variance in Salinity is explained by Trial. (18% when Control not included)
39% of variance in Temperature is explained by Trial.

#### Output Data
```{r env-fact-vary-trial-kruskal-output-data}
# stats results
#write_xlsx(kruskal_all, "OutputData/ConfoundingFactors/ConfoundingFactorTesting-KruskalWallis-InitialResults.xlsx")

# effect size results
#write_xlsx(kruskal_all_effsize, "OutputData/ConfoundingFactors/ConfoundingFactorTesting-KruskalWallis-EffectSize.xlsx")

```


### Dunn's Test of Multiple Comparisons
```{r env-fact-vary-trial-all-dunn}
# dunn test for nonparametric pairwise multiple comparisons of ranked data
dunn_all <- dfAllEnvVar_long %>%
  group_by(variable_names) %>%
  dunn_test(variable_values ~ Trial, p.adjust.method = "bonferroni", detailed = TRUE) %>%
  select(-p, -method) %>%
  mutate(p.adj = round(p.adj, 4))

# make excel file of results
# write_xlsx(dunn_all, "OutputData/ConfoundingFactors/ConfoundingFactorTesting-Dunn.xlsx")

```

Significant variables from Dunn test: 
- Alkalinity
- Ammonia
- Calcium
- Chlorophyll
- ORP
- pH
- Phycoerytherin
- Salinity 
- Temperature

See ConfoundingFactorTesting-Dunn.xlsx for details (estimates are mean ranks difference rather than difference in means/medians)

## Do the environmental variables influence the response variables? 
### Data Setup
Since the environmental data is a time series, but the coral data isn't, and we want to compare them directly, we'll take the mean of the environmental data for each trial but leave the coral data alone. This will give us a dataframe that looks like different values for the coral data with the same environmental value for each coral data point for each trial. `View(dfCoralRateAllEnvMean)` to see what that looks like. 
#### Experimental
```{r env-fact-infl-data-setup-experimental}
# mean environmental variable per trial
dfAllEnvVar_mean_exp <- dfAllEnvVar %>%
  filter(Treatment != "Control") %>%
  group_by(Treatment, Trial) %>%
  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)))

# combine with coral data 
dfCoralRateAllEnvMean_exp <- dfCoral_rates %>%
  filter(Treatment != "Control") %>%
  left_join(dfAllEnvVar_mean_exp) %>%
  mutate(Trial = factor(Trial, levels = c("PBR 1", "PBC 1", "PBR 2", "PBR 3", "PBC 2", "PBC 3")))

```

#### Control
```{r env-fact-infl-data-setup-experimental}
# mean environmental variable per trial
dfAllEnvVar_mean_control <- dfAllEnvVar %>%
  filter(Treatment == "Control") %>%
  group_by(Treatment, Trial) %>%
  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)))

# combine with coral data 
dfCoralRateAllEnvMean_control <- dfCoral_rates %>%
  filter(Treatment == "Control") %>%
  left_join(dfAllEnvVar_mean_control) %>%
  mutate(Trial = factor(Trial, levels = c("Control")))

```

### Linear Models
#### Regular Linear Regression
##### Define Functions
Make functions for linear models using dependent variables and grouped by environmental variable 
```{r lm-functions}
# subracted
# weight
func_lm_weight <- function(df) {
  lm(weight_g_weekly ~ var_values, data = df)
}

# surface area
func_lm_surface_area <- function(df) {
  lm(surface_area_cm3_weekly ~ var_values, data = df)
}
# rgb
func_lm_rgb <- function(df) {
  lm(rgb_weekly ~ var_values, data = df)
}

```

##### Regressions
###### Experimental
```{r lm-regressions-experimental}
# make nested list of pivoted dataframes. nested by dependent variable
lm_data_list_exp <- dfCoralRateAllEnvMean_exp %>%
  pivot_longer(cols = 8:22,
               names_to = "var_names",
               values_to = "var_values") %>%
  group_by(var_names) %>%
  nest()

# apply lm functions to their respective variable in the nested list
lm_data_list_exp <- lm_data_list_exp %>%
  mutate(lm_weight = map(data, func_lm_weight),
         lm_surface_area = map(data, func_lm_surface_area),
         lm_rgb = map(data, func_lm_rgb))

# Grab Results
#### weight ####
# empty list to put lm weight results in
lm_results_weight_exp <- data.frame(environmental_variable = lm_data_list_exp$var_names,
                                p_value = NA,
                                f_stat = NA,
                                r_squared = NA, 
                                estimate = NA)
# loop through length of lm_results_weight_exp and grab p-value, f statistic, and r^2 value for each model from lm_data_list_exp 
for (i in 1:length(lm_data_list_exp$var_names)) {
  # p value for i 
  lm_results_weight_exp$p_value[i] <- summary(lm_data_list_exp$lm_weight[[i]])$coefficients[2,4]
  # f stat for i 
  lm_results_weight_exp$f_stat[i] <- summary(lm_data_list_exp$lm_weight[[i]])$fstatistic[1]
  # r^2 for i
  lm_results_weight_exp$r_squared[i] <- summary(lm_data_list_exp$lm_weight[[i]])$r.squared
  # estimate for i
  lm_results_weight_exp$estimate[i] <- summary(lm_data_list_exp$lm_weight[[i]])$coefficients[2]
}
# round values to 4th decimal place 
lm_results_weight_exp <- lm_results_weight_exp %>%
  mutate(across(where(is.numeric), ~round(.x, digits = 4)))

#### surface area ####
# empty list to put lm surface area results in
lm_results_surface_area_exp <- data.frame(environmental_variable = lm_data_list_exp$var_names,
                                      p_value = NA,
                                      f_stat = NA,
                                      r_squared = NA,
                                      estimate = NA)
# loop through length of lm_results_surface_area_exp and grab p-value, f statistic, and r^2 value for each model from lm_data_list_exp 
for (i in 1:length(lm_data_list_exp$var_names)) {
  # p value for i 
  lm_results_surface_area_exp$p_value[i] <- summary(lm_data_list_exp$lm_surface_area[[i]])$coefficients[2,4]
  # f stat for i 
  lm_results_surface_area_exp$f_stat[i] <- summary(lm_data_list_exp$lm_surface_area[[i]])$fstatistic[1]
  # r^2 for i
  lm_results_surface_area_exp$r_squared[i] <- summary(lm_data_list_exp$lm_surface_area[[i]])$r.squared
  # estimate for i
  lm_results_surface_area_exp$estimate[i] <- summary(lm_data_list_exp$lm_surface_area[[i]])$coefficients[2]
}
# round values to 4th decimal place 
lm_results_surface_area_exp <- lm_results_surface_area_exp %>%
  mutate(across(where(is.numeric), ~round(.x, digits = 4)))

#### rgb ####
# empty list to put lm rgb results in
lm_results_rgb_exp <- data.frame(environmental_variable = lm_data_list_exp$var_names,
                             p_value = NA,
                             f_stat = NA,
                             r_squared = NA,
                             estimate = NA)
# loop through length of lm_results_rgb_exp and grab p-value, f statistic, and r^2 value for each model from lm_data_list_exp 
for (i in 1:length(lm_data_list_exp$var_names)) {
  # p value for i 
  lm_results_rgb_exp$p_value[i] <- summary(lm_data_list_exp$lm_rgb[[i]])$coefficients[2,4]
  # f stat for i 
  lm_results_rgb_exp$f_stat[i] <- summary(lm_data_list_exp$lm_rgb[[i]])$fstatistic[1]
  # r^2 for i
  lm_results_rgb_exp$r_squared[i] <- summary(lm_data_list_exp$lm_rgb[[i]])$r.squared
  # estimate for i
  lm_results_rgb_exp$estimate[i] <- summary(lm_data_list_exp$lm_rgb[[i]])$coefficients[2]
}
# round values to 4th decimal place 
lm_results_rgb_exp <- lm_results_rgb_exp %>%
  mutate(across(where(is.numeric), ~round(.x, digits = 4)))

#### export results ####
# prep lm dataframes to combine
# weight
lm_results_weight_exp <- lm_results_weight_exp %>%
  add_column(dependent_variable = "weight_g_weekly", 
             .before = "environmental_variable") 
# surface area
lm_results_surface_area_exp <- lm_results_surface_area_exp %>%
  add_column(dependent_variable = "surface_area_cm3_weekly",
             .before = "environmental_variable")
# rgb
lm_results_rgb_exp <- lm_results_rgb_exp %>%
  add_column(dependent_variable = "rgb_weekly",
             .before = "environmental_variable")
# combine dataframes
lm_results_exp <- rbind(lm_results_weight_exp, lm_results_surface_area_exp, lm_results_rgb_exp) %>%
  filter(p_value < 0.05) %>%
  mutate(r_squared = round(r_squared, digits = 2))


# write excel file
# write_xlsx(lm_results_exp, "OutputData/ConfoundingFactors/ConfoundingFactorTesting-LinearRegressions-Experimental.xlsx")

```
Significant Results: 
Weight: 
- ORP: r^2 = 0.14
- Phycoerythrin: r^2 = 0.10
- Nitrate MR: r^2 = 0.19
- Silica: r^2 = 0.18
- Calcium: r^2 = 0.12
Surface area: 
- Temperature
- Salinity
- pH: r^2 = 0.11
- ORP
- Nitrate MR: r^2 = 0.13
- Silica
- Ammonia
RGB: 
- Salinity 
- ORP
- Phycoerythrin
- Alkalinity
- Calcium

Quite a few variables have a significant linear relationship with weight, but none of them have a higher r^2 than 0.20, meaning that the relationships are fairly weak. 

###### Control 
```{r lm-regressions-experimental}
# make nested list of pivoted dataframes. nested by dependent variable
lm_data_list_control <- dfCoralRateAllEnvMean_control %>%
  pivot_longer(cols = 8:22,
               names_to = "var_names",
               values_to = "var_values") %>%
  group_by(var_names) %>%
  nest()

# apply lm functions to their respective variable in the nested list
lm_data_list_control <- lm_data_list_control %>%
  mutate(lm_weight = map(data, func_lm_weight),
         lm_surface_area = map(data, func_lm_surface_area),
         lm_rgb = map(data, func_lm_rgb))

# Grab Results
#### weight ####
# empty list to put lm weight results in
lm_results_weight_control <- data.frame(environmental_variable = lm_data_list_control$var_names,
                                p_value = NA,
                                f_stat = NA,
                                r_squared = NA, 
                                estimate = NA)
# loop through length of lm_results_weight_control and grab p-value, f statistic, and r^2 value for each model from lm_data_list_control 
for (i in 1:length(lm_data_list_control$var_names)) {
  # p value for i 
  lm_results_weight_control$p_value[i] <- summary(lm_data_list_control$lm_weight[[i]])$coefficients[2,4]
  # f stat for i 
  lm_results_weight_control$f_stat[i] <- summary(lm_data_list_control$lm_weight[[i]])$fstatistic[1]
  # r^2 for i
  lm_results_weight_control$r_squared[i] <- summary(lm_data_list_control$lm_weight[[i]])$r.squared
  # estimate for i
  lm_results_weight_control$estimate[i] <- summary(lm_data_list_control$lm_weight[[i]])$coefficients[2]
}
# round values to 4th decimal place 
lm_results_weight_control <- lm_results_weight_control %>%
  mutate(across(where(is.numeric), ~round(.x, digits = 4)))

#### surface area ####
# empty list to put lm surface area results in
lm_results_surface_area_control <- data.frame(environmental_variable = lm_data_list_control$var_names,
                                      p_value = NA,
                                      f_stat = NA,
                                      r_squared = NA,
                                      estimate = NA)
# loop through length of lm_results_surface_area_control and grab p-value, f statistic, and r^2 value for each model from lm_data_list_control 
for (i in 1:length(lm_data_list_control$var_names)) {
  # p value for i 
  lm_results_surface_area_control$p_value[i] <- summary(lm_data_list_control$lm_surface_area[[i]])$coefficients[2,4]
  # f stat for i 
  lm_results_surface_area_control$f_stat[i] <- summary(lm_data_list_control$lm_surface_area[[i]])$fstatistic[1]
  # r^2 for i
  lm_results_surface_area_control$r_squared[i] <- summary(lm_data_list_control$lm_surface_area[[i]])$r.squared
  # estimate for i
  lm_results_surface_area_control$estimate[i] <- summary(lm_data_list_control$lm_surface_area[[i]])$coefficients[2]
}
# round values to 4th decimal place 
lm_results_surface_area_control <- lm_results_surface_area_control %>%
  mutate(across(where(is.numeric), ~round(.x, digits = 4)))

#### rgb ####
# empty list to put lm rgb results in
lm_results_rgb_control <- data.frame(environmental_variable = lm_data_list_control$var_names,
                             p_value = NA,
                             f_stat = NA,
                             r_squared = NA,
                             estimate = NA)
# loop through length of lm_results_rgb_control and grab p-value, f statistic, and r^2 value for each model from lm_data_list_control 
for (i in 1:length(lm_data_list_control$var_names)) {
  # p value for i 
  lm_results_rgb_control$p_value[i] <- summary(lm_data_list_control$lm_rgb[[i]])$coefficients[2,4]
  # f stat for i 
  lm_results_rgb_control$f_stat[i] <- summary(lm_data_list_control$lm_rgb[[i]])$fstatistic[1]
  # r^2 for i
  lm_results_rgb_control$r_squared[i] <- summary(lm_data_list_control$lm_rgb[[i]])$r.squared
  # estimate for i
  lm_results_rgb_control$estimate[i] <- summary(lm_data_list_control$lm_rgb[[i]])$coefficients[2]
}
# round values to 4th decimal place 
lm_results_rgb_control <- lm_results_rgb_control %>%
  mutate(across(where(is.numeric), ~round(.x, digits = 4)))

#### export results ####
# prep lm dataframes to combine
# weight
lm_results_weight_control <- lm_results_weight_control %>%
  add_column(dependent_variable = "weight_g_weekly", 
             .before = "environmental_variable") 
# surface area
lm_results_surface_area_control <- lm_results_surface_area_control %>%
  add_column(dependent_variable = "surface_area_cm3_weekly",
             .before = "environmental_variable")
# rgb
lm_results_rgb_control <- lm_results_rgb_control %>%
  add_column(dependent_variable = "rgb_weekly",
             .before = "environmental_variable")
# combine dataframes
lm_results_control <- rbind(lm_results_weight_control, lm_results_surface_area_control, lm_results_rgb_control) %>%
  filter(p_value < 0.05) %>%
  mutate(r_squared = round(r_squared, digits = 2))


# write excel file
# write_xlsx(lm_results_all, "OutputData/ConfoundingFactors/ConfoundingFactorTesting-LinearRegressions-Control.xlsx")

```

#### Weighted Least Squares Regression
This is a form of linear regression that places more weight on the observations with smaller error variance. To see which variables require this test, we will first run a Breusch-Pagan test to see if there is homoscedasticity in the residuals. If the test has a p-value < 0.05, then there is homoscedasticity. If homoscedasticity is present, then we will run the weighted least squares regression. Otherwise, the unweighted linear regression is fine. 
Source: https://www.statology.org/weighted-least-squares-in-r/
##### Define Functions
```{r wls-functions}
# make functions for weighted least squares linear models using dependent vars
# weight
func_wls_weight <- function(df) {
  lm(formula = weight_g_weekly ~ var_values, data = df, 
     weights = 1/lm(abs(resid) ~ pred)$fitted.values^2)
}

# wls_list_weight <- lm_resids_weight %>%
#  mutate(wls = map(data, func_wls_weight))

# surface area
func_wls_surface_area <- function(df) {
  lm(formula = surface_area_cm3_weekly ~ var_values, data = df, 
     weights = 1/lm(abs(resid) ~ pred)$fitted.values^2)
}
# rgb
func_wls_rgb <- function(df) {
  lm(formula = rgb_weekly ~ var_values, data = df, 
     weights = 1/lm(abs(resid) ~ pred)$fitted.values^2)
}

```

##### Prepare Data
```{r wls-data-prep}
# add residuals for each model 
lm_data_list <- lm_data_list %>%
  mutate(resids_weight = map2(data, lm_weight, add_residuals),
         resids_surface_area = map2(data, lm_surface_area, add_residuals),
         resids_rgb = map2(data, lm_rgb, add_residuals),
         fit_vals_weight = map2(data, lm_weight, add_predictions),
         fit_vals_surface_area = map2(data, lm_surface_area, add_predictions),
         fit_vals_rgb = map2(data, lm_rgb, add_predictions))

# unnest and separate into dataframes based on dependent variable then re-nest by environmental variable to prep for WLS regressions in next step
# weight
lm_resids_weight <- unnest(lm_data_list, resids_weight, fit_vals_weight) %>%
  select(var_names, weight_g_weekly, var_values, resid, pred) %>%
  na.omit() %>%
  group_by(var_names) %>%
  nest()

# surface area
lm_resids_surface_area <- unnest(lm_data_list, resids_surface_area, fit_vals_surface_area) %>%
  select(var_names, surface_area_cm3_weekly, var_values, resid, pred) %>%
  na.omit() %>%
  group_by(var_names) %>%
  nest()

# rgb
lm_resids_rgb <- unnest(lm_data_list, resids_rgb, fit_vals_rgb) %>%
  select(var_names, rgb_weekly, var_values, resid, pred) %>%
  na.omit() %>%
  group_by(var_names) %>%
  nest()

```


##### Regressions
Perform wls regressions for each environmental variable in each dependent variable and output results: estimate, degrees of freedom (df), residual degrees of freedom (rdf), standard error (se), residual standard error (rse), F statistic (f_stat), r^2, p-value.
```{r wls}
#### weight ####
wls_list_weight <- lm_resids_weight %>%
  mutate(wls = map(data, func_wls_weight))

# empty list to put wls weight results in
wls_results_weight <- data.frame(environmental_variable = wls_list_weight$var_names,
                                 estimate = NA,
                                 df = NA,
                                 rdf = NA,
                                 se = NA,
                                 rse = NA,
                                 f_stat = NA,
                                 r_squared = NA,
                                 p_value = NA
                                 )
# loop through length of wls_results_weight and grab p-value, f statistic, and r^2 value for each model from wls_list_weight 
for (i in 1:length(wls_list_weight$var_names)) {
  # estimate for i
  wls_results_weight$estimate[i] <- summary(wls_list_weight$wls[[i]])$coefficients[2]
  # degrees of freedom for i
  wls_results_weight$df[i] <- summary(wls_list_weight$wls[[i]])$df[1]
  # residual degrees of freedom for i 
  wls_results_weight$rdf[i] <- summary(wls_list_weight$wls[[i]])$df[2]
  # standard error for i
  wls_results_weight$se[i] <-summary(wls_list_weight$wls[[i]])$coefficients[2, "Std. Error"]
  # residual standard error for i
  wls_results_weight$rse[i] <-summary(wls_list_weight$wls[[i]])$sigma
  # f stat for i 
  wls_results_weight$f_stat[i] <- summary(wls_list_weight$wls[[i]])$fstatistic[1]
  # r^2 for i
  wls_results_weight$r_squared[i] <- summary(wls_list_weight$wls[[i]])$r.squared
  # p value for i 
  wls_results_weight$p_value[i] <- summary(wls_list_weight$wls[[i]])$coefficients[2,4]
}
# round values to 4th decimal place 
wls_results_weight <- wls_results_weight %>%
  mutate(across(where(is.numeric), ~round(.x, digits = 4)))


#### surface area ####
wls_list_surface_area <- lm_resids_surface_area %>%
  mutate(wls = map(data, func_wls_surface_area))

# empty list to put wls weight results in
wls_results_surface_area <- data.frame(environmental_variable = wls_list_surface_area$var_names,
                                       estimate = NA,
                                       df = NA,
                                       rdf = NA,
                                       se = NA,
                                       rse = NA,
                                       f_stat = NA,
                                       r_squared = NA,
                                       p_value = NA)
# loop through length of wls_results_surface_area and grab p-value, f statistic, and r^2 value for each model from wls_list_surface_area 
for (i in 1:length(wls_list_surface_area$var_names)) {
  # estimate for i
  wls_results_surface_area$estimate[i] <- summary(wls_list_surface_area$wls[[i]])$coefficients[2]
  # degrees of freedom for i
  wls_results_surface_area$df[i] <- summary(wls_list_surface_area$wls[[i]])$df[1]
  # residual degrees of freedom for i 
  wls_results_surface_area$rdf[i] <- summary(wls_list_surface_area$wls[[i]])$df[2]
  # standard error for i
  wls_results_surface_area$se[i] <-summary(wls_list_surface_area$wls[[i]])$coefficients[2, "Std. Error"]
  # residual standard error for i
  wls_results_surface_area$rse[i] <-summary(wls_list_surface_area$wls[[i]])$sigma
  # f stat for i 
  wls_results_surface_area$f_stat[i] <- summary(wls_list_surface_area$wls[[i]])$fstatistic[1]
  # r^2 for i
  wls_results_surface_area$r_squared[i] <- summary(wls_list_surface_area$wls[[i]])$r.squared
  # p value for i 
  wls_results_surface_area$p_value[i] <- summary(wls_list_surface_area$wls[[i]])$coefficients[2,4]
}
# round values to 4th decimal place 
wls_results_surface_area <- wls_results_surface_area %>%
  mutate(across(where(is.numeric), ~round(.x, digits = 4)))


#### rgb ####
wls_list_rgb <- lm_resids_rgb %>%
  mutate(wls = map(data, func_wls_rgb))

# empty list to put wls weight results in
wls_results_rgb <- data.frame(environmental_variable = wls_list_rgb$var_names,
                              estimate = NA,
                              df = NA,
                              rdf = NA,
                              se = NA,
                              rse = NA,
                              f_stat = NA,
                              r_squared = NA,
                              p_value = NA)
# loop through length of wls_results_rgb and grab p-value, f statistic, and r^2 value for each model from wls_list_rgb 
for (i in 1:length(wls_list_rgb$var_names)) {
  # estimate for i
  wls_results_rgb$estimate[i] <- summary(wls_list_rgb$wls[[i]])$coefficients[2]
  # degrees of freedom for i
  wls_results_rgb$df[i] <- summary(wls_list_rgb$wls[[i]])$df[1]
  # residual degrees of freedom for i 
  wls_results_rgb$rdf[i] <- summary(wls_list_rgb$wls[[i]])$df[2]
  # standard error for i
  wls_results_rgb$se[i] <-summary(wls_list_rgb$wls[[i]])$coefficients[2, "Std. Error"]
  # residual standard error for i
  wls_results_rgb$rse[i] <-summary(wls_list_rgb$wls[[i]])$sigma
  # f stat for i 
  wls_results_rgb$f_stat[i] <- summary(wls_list_rgb$wls[[i]])$fstatistic[1]
  # r^2 for i
  wls_results_rgb$r_squared[i] <- summary(wls_list_rgb$wls[[i]])$r.squared
  # p value for i 
  wls_results_rgb$p_value[i] <- summary(wls_list_rgb$wls[[i]])$coefficients[2,4]
}
# round values to 4th decimal place 
wls_results_rgb <- wls_results_rgb %>%
  mutate(across(where(is.numeric), ~round(.x, digits = 4)))

#### Export Results ####
# prep lm dataframes to combine
# weight
wls_results_weight <- wls_results_weight %>%
  add_column(dependent_variable = "weight_g_weekly", 
             .before = "environmental_variable") 
# surface area
wls_results_surface_area <- wls_results_surface_area %>%
  add_column(dependent_variable = "surface_area_cm3_weekly",
             .before = "environmental_variable")
# rgb
wls_results_rgb <- wls_results_rgb %>%
  add_column(dependent_variable = "rgb_weekly",
             .before = "environmental_variable")
# combine dataframes
wls_results_all <- rbind(wls_results_weight, wls_results_surface_area, wls_results_rgb) %>%
  filter(p_value < 0.05)

# write excel file
# write_xlsx(wls_results_all, "OutputData/ConfoundingFactors/ConfoundingFactorTesting-WeightedLeastSquareRegressions.xlsx")


```
Significant Results: 
Weight: 
- Temperature
- pH
- ORP
- Phycoerytherin 
- Nitrate MR: r^2 = 0.18
- Silica: r^2 = 0.18
- Iron
- Calcium 
Surface Area: r^2 all 0.10 or less 
- Temperature
- Nitrate 
- Alkalinity 
- Silica
- Ammonia
RGB: r^2 all 0.10 or less 
- Temperature
- Salinity 
- pH
- Silica
- Iron
- Calcium

Quite a few variables have a significant linear relationship with weight, but none of them have a higher r^2 than 0.20, meaning that the relationships are fairly weak. 


#### Weighted Multiple Linear Regression
Instead of looking at how each environmental variable individually affected the dependent varaibles, a multiple linear regression will add in the effects from multiple environmental variables. 

##### Prepare Data
The weighted data will only work if every variable has the same amount of data. DLI, Alkalinity, Silica, and Iron all have missing values. DLI is missing Control, Alkalinity is missing PBR 1, and Silica and Iron are missing PBC 2, PBC 3, and PBR 3. 
```{r wls-mult-data-prep}
# remove Control treatment bc dli doesnt have a control unfortunately 
dfCoralRateAllEnvMean_ex <- dfCoralRateAllEnvMean %>%
  filter(Treatment != "Control")

```

##### Regression
```{r wls-mult}
# model1 <- lm(Total.Wasps ~ Temperature.Outside + Temperature.Inside + HourAdjusted, data = wasps2) 

# # Sample multiple regression data frame
# df <- data.frame(y, x1, x2)
# 
# # Step 1: Fit ordinary least squares model
# ols_fit <- lm(y ~ x1 + x2, data = df)
# 
# # Step 2: Calculate weights based on residual variance
# wts <- 1 / fitted(lm(abs(residuals(ols_fit)) ~ fitted(ols_fit)))^2
# 
# # Step 3: Fit weighted least squares model
# wls_fit <- lm(y ~ x1 + x2, data = df, weights = wts)
# 
# # View results
# summary(wls_fit)


# weight 
# fit ordinary least squares model 
## all variables 
ols_mult_fit <- lm(weight_g_weekly ~ `Temperature (ºC)` + `Salinity (PSU)` + pH +
                     `ORP (mV)` + `Chlorophyll (RFU)` + `Phycoerythrin (RFU)` + 
                     DLI + Turbidity + Phosphate + `Nitrate MR` + Ammonia + Calcium,
                   data = dfCoralRateAllEnvMean)

# calculate weights based on residual variance 
wts <- 1/ fitted(lm(abs(residuals(ols_mult_fit)) ~ fitted(ols_mult_fit)))^2

# fit weighted least squares model
wls_mult_fit <- lm(weight_g_weekly ~ `Temperature (ºC)` + `Salinity (PSU)` + pH +
                     `ORP (mV)` + `Chlorophyll (RFU)` + `Phycoerythrin (RFU)` + 
                     DLI + Turbidity + Phosphate + `Nitrate MR` + Ammonia + Calcium,
                   data = dfCoralRateAllEnvMean,
                   weights = wts)

summary(wls_mult_fit)


```



#### Summary of Linear and Weighted Least Squares Regression Models
Quite a few variables have a significant linear relationship with weight, but none of them have a higher r^2 than 0.20, meaning that the relationships are fairly weak. 

# Plots
## Weight 
### Scatter Plots
#### Significant
```{r plots-weight}
# temperature 
scatter_weight_temp <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Temperature (ºC)`,
             y = weight_g_weekly, 
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (weight_g_weekly - sd(weight_g_weekly, na.rm = TRUE)),
                    ymax = (weight_g_weekly + sd(weight_g_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Weight (g) vs Mean Temperature (ºC) Per Trial") +
  ylab("Weight (g)") +
  xlab("Temperature (ºC)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_weight_temp

# pH
scatter_weight_ph <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = pH,
             y = weight_g_weekly, 
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (weight_g_weekly - sd(weight_g_weekly, na.rm = TRUE)),
                    ymax = (weight_g_weekly + sd(weight_g_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Weight (g) vs Mean pH Per Trial") +
  ylab("Weight (g)") +
  xlab("pH") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_weight_ph

# ORP
scatter_weight_orp <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `ORP (mV)`,
             y = weight_g_weekly, 
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (weight_g_weekly - sd(weight_g_weekly, na.rm = TRUE)),
                    ymax = (weight_g_weekly + sd(weight_g_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Weight (g) vs Mean ORP (mV) Per Trial") +
  ylab("Weight (g)") +
  xlab("ORP (mV)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_weight_orp

# Phycoerytherin
scatter_weight_phyco <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Phycoerythrin (RFU)`,
             y = weight_g_weekly, 
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (weight_g_weekly - sd(weight_g_weekly, na.rm = TRUE)),
                    ymax = (weight_g_weekly + sd(weight_g_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Weight (g) vs Mean Phycoerythrin (RFU) Per Trial") +
  ylab("Weight (g)") +
  xlab("Phycoerythrin (RFU)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_weight_phyco

# Nitrate MR
scatter_weight_nit_mr <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Nitrate MR`,
             y = weight_g_weekly, 
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (weight_g_weekly - sd(weight_g_weekly, na.rm = TRUE)),
                    ymax = (weight_g_weekly + sd(weight_g_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Weight (g) vs Mean Nitrate MR (mg/L) Per Trial") +
  ylab("Weight (g)") +
  xlab("Nitrate MR (mg/L)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_weight_nit_mr

# silica
scatter_weight_silica <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Silica`,
             y = weight_g_weekly, 
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (weight_g_weekly - sd(weight_g_weekly, na.rm = TRUE)),
                    ymax = (weight_g_weekly + sd(weight_g_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Weight (g) vs Mean Silica (mg/L) Per Trial") +
  ylab("Weight (g)") +
  xlab("Silica (mg/L)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_weight_silica

# iron
scatter_weight_iron <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Iron`,
             y = weight_g_weekly, 
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (weight_g_weekly - sd(weight_g_weekly, na.rm = TRUE)),
                    ymax = (weight_g_weekly + sd(weight_g_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Weight (g) vs Mean Iron (mg/L) Per Trial") +
  ylab("Weight (g)") +
  xlab("Iron (mg/L)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_weight_iron

# calcium
scatter_weight_calcium <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Calcium`,
             y = weight_g_weekly, 
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (weight_g_weekly - sd(weight_g_weekly, na.rm = TRUE)),
                    ymax = (weight_g_weekly + sd(weight_g_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Weight (g) vs Mean Calcium (mg/L) Per Trial") +
  ylab("Weight (g)") +
  xlab("Calcium (mg/L)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_weight_calcium

```

#### Not Significant
```{r plots-weight-scatter-not-sig}
# salinity
scatter_weight_sal <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Salinity (PSU)`,
             y = weight_g_weekly,
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (weight_g_weekly - sd(weight_g_weekly, na.rm = TRUE)),
                    ymax = (weight_g_weekly + sd(weight_g_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Weight (g) vs Mean Salinity Per Trial") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_weight_sal

# Chlorophyll
scatter_weight_chlor <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Chlorophyll (RFU)`,
             y = weight_g_weekly,
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (weight_g_weekly - sd(weight_g_weekly, na.rm = TRUE)),
                    ymax = (weight_g_weekly + sd(weight_g_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Weight (g) vs Mean Chlorophyll (RFU) Per Trial") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_weight_chlor

# DLI
scatter_weight_dli <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = DLI,
             y = weight_g_weekly,
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (weight_g_weekly - sd(weight_g_weekly, na.rm = TRUE)),
                    ymax = (weight_g_weekly + sd(weight_g_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Weight (g) vs Mean DLI (mol*m^-2*s^-1) Per Trial") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_weight_dli

# Turbidity
scatter_weight_turb <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = Turbidity,
             y = weight_g_weekly,
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (weight_g_weekly - sd(weight_g_weekly, na.rm = TRUE)),
                    ymax = (weight_g_weekly + sd(weight_g_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Weight (g) vs Mean Turbidity (TAU) Per Trial") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_weight_turb

# Phosphate
scatter_weight_phos <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = Phosphate,
             y = weight_g_weekly,
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (weight_g_weekly - sd(weight_g_weekly, na.rm = TRUE)),
                    ymax = (weight_g_weekly + sd(weight_g_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Weight (g) vs Mean Phosphate (mg/L) Per Trial") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_weight_phos

# Alkalinity
scatter_weight_alkalinity <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = Alkalinity,
             y = weight_g_weekly,
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (weight_g_weekly - sd(weight_g_weekly, na.rm = TRUE)),
                    ymax = (weight_g_weekly + sd(weight_g_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Weight (g) vs Mean Alkalinity (mg/L) Per Trial") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_weight_alkalinity

# Ammonia
scatter_weight_ammonia <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = Ammonia,
             y = weight_g_weekly,
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (weight_g_weekly - sd(weight_g_weekly, na.rm = TRUE)),
                    ymax = (weight_g_weekly + sd(weight_g_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Weight (g) vs Mean Ammonia (mg/L) Per Trial") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_weight_ammonia

```

### Box Plots
#### Significant
```{r plots-weight-boxplot-sig}
# temp
box_weight_temp <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Temperature (ºC)`,
             y = weight_g_weekly, 
             color = Trial)) +
  geom_boxplot() +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Weight (g) vs Mean Temperature (ºC) Per Trial") +
  ylab("Weight (g)") +
  xlab("Temperature (ºC)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
box_weight_temp

# pH 
box_weight_pH <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = pH,
             y = weight_g_weekly, 
             color = Trial)) +
  geom_boxplot() +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Weight (g) vs Mean pH Per Trial") +
  ylab("Weight (g)") +
  xlab("pH") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
box_weight_pH

# orp 
box_weight_orp <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `ORP (mV)`,
             y = weight_g_weekly, 
             color = Trial)) +
  geom_boxplot() +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Weight (g) vs Mean ORP (mV) Per Trial") +
  ylab("Weight (g)") +
  xlab("ORP (mV)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
box_weight_orp

# phycoerytherin
box_weight_phyco <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Phycoerythrin (RFU)`,
             y = weight_g_weekly, 
             color = Trial)) +
  geom_boxplot() +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Weight (g) vs Mean Phycoerythrin (RFU) Per Trial") +
  ylab("Weight (g)") +
  xlab("Phycoerythrin (RFU)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
box_weight_phyco

# nitrate mr
box_weight_nit_mr <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Nitrate MR`,
             y = weight_g_weekly, 
             color = Trial)) +
  geom_boxplot() +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Weight (g) vs Mean Nitrate MR Per Trial") +
  ylab("Weight (g)") +
  xlab("Nitrate MR") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
box_weight_nit_mr

# silica
box_weight_silica <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Silica`,
             y = weight_g_weekly, 
             color = Trial)) +
  geom_boxplot() +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Weight (g) vs Mean Silica Per Trial") +
  ylab("Weight (g)") +
  xlab("Silica (mg/L)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
box_weight_silica

# iron
box_weight_iron <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Iron`,
             y = weight_g_weekly, 
             color = Trial)) +
  geom_boxplot() +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Weight (g) vs Mean Iron Per Trial") +
  ylab("Weight (g)") +
  xlab("Iron (mg/L)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
box_weight_iron

# calcium
box_weight_calc <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Calcium`,
             y = weight_g_weekly, 
             color = Trial)) +
  geom_boxplot() +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Weight (g) vs Mean Calcium Per Trial") +
  ylab("Weight (g)") +
  xlab("Calcium (mg/L)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
box_weight_calc

```


## Surface Area
### Scatter Plots
#### Significant
```{r scatter-surface-area-sig}
# temperature
scatter_sur_area_temp <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Temperature (ºC)`,
             y = surface_area_cm3_weekly, 
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (surface_area_cm3_weekly - sd(surface_area_cm3_weekly, na.rm = TRUE)),
                    ymax = (surface_area_cm3_weekly + sd(surface_area_cm3_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Surface Area (cm^3) vs Mean Temperature (ºC) Per Trial") +
  ylab("Surface Area (cm^3)") +
  xlab("Temperature (ºC)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_sur_area_temp

# Nitrate MR
scatter_sur_area_nit_mr <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Nitrate MR`,
             y = surface_area_cm3_weekly, 
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (surface_area_cm3_weekly - sd(surface_area_cm3_weekly, na.rm = TRUE)),
                    ymax = (surface_area_cm3_weekly + sd(surface_area_cm3_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3")) +
  ggtitle("Surface Area (cm^3) vs Mean Nitrate MR (mg/L) Per Trial") +
  ylab("Surface Area (cm^3)") +
  xlab("Nitrate MR (mg/L)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_sur_area_nit_mr

# alkalinity
scatter_sur_area_alk <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Alkalinity`,
             y = surface_area_cm3_weekly, 
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (surface_area_cm3_weekly - sd(surface_area_cm3_weekly, na.rm = TRUE)),
                    ymax = (surface_area_cm3_weekly + sd(surface_area_cm3_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3")) +
  ggtitle("Surface Area (cm^3) vs Mean Alkalinity (mg/L) Per Trial") +
  ylab("Surface Area (cm^3)") +
  xlab("Alkalinity (mg/L)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_sur_area_alk

# silica
scatter_sur_area_silica <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Silica`,
             y = surface_area_cm3_weekly, 
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (surface_area_cm3_weekly - sd(surface_area_cm3_weekly, na.rm = TRUE)),
                    ymax = (surface_area_cm3_weekly + sd(surface_area_cm3_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3")) +
  ggtitle("Surface Area (cm^3) vs Mean Silica (mg/L) Per Trial") +
  ylab("Surface Area (cm^3)") +
  xlab("Silica (mg/L)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_sur_area_silica

# ammonia 
scatter_sur_area_ammo <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Ammonia`,
             y = surface_area_cm3_weekly, 
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (surface_area_cm3_weekly - sd(surface_area_cm3_weekly, na.rm = TRUE)),
                    ymax = (surface_area_cm3_weekly + sd(surface_area_cm3_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3")) +
  ggtitle("Surface Area (cm^3) vs Mean Ammonia (mg/L) Per Trial") +
  ylab("Surface Area (cm^3)") +
  xlab("Ammonia (mg/L)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_sur_area_ammo

```

### Box Plots
#### Significant
```{r boxplots-sur-area-sig}
# temperature 
box_sur_area_temp <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Temperature (ºC)`,
             y = surface_area_cm3_weekly, 
             color = Trial)) +
  geom_boxplot() +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Surface Area (cm^3) vs Mean Temperature (ºC) Per Trial") +
  ylab("Surface Area (cm^3)") +
  xlab("Temperature (ºC)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
box_sur_area_temp

# nitrate
box_sur_area_nit <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Nitrate MR`,
             y = surface_area_cm3_weekly, 
             color = Trial)) +
  geom_boxplot() +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Surface Area (cm^3) vs Mean Nitrate (mg/L) Per Trial") +
  ylab("Surface Area (cm^3)") +
  xlab("Nitrate (mg/L)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
box_sur_area_nit

# alkalinity
box_sur_area_alk <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Alkalinity`,
             y = surface_area_cm3_weekly, 
             color = Trial)) +
  geom_boxplot() +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Surface Area (cm^3) vs Mean Alkalinity (mg/L) Per Trial") +
  ylab("Surface Area (cm^3)") +
  xlab("Alkalinity (mg/L)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
box_sur_area_alk

# silica
box_sur_area_silica <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Silica`,
             y = surface_area_cm3_weekly, 
             color = Trial)) +
  geom_boxplot() +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Surface Area (cm^3) vs Mean Silica (mg/L) Per Trial") +
  ylab("Surface Area (cm^3)") +
  xlab("Silica (mg/L)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
box_sur_area_silica

# ammonia
box_sur_area_ammo <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Ammonia`,
             y = surface_area_cm3_weekly, 
             color = Trial)) +
  geom_boxplot() +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Surface Area (cm^3) vs Mean Ammonia (mg/L) Per Trial") +
  ylab("Surface Area (cm^3)") +
  xlab("Ammonia (mg/L)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
box_sur_area_ammo


```


## RGB
### Scatter Plots
#### Significant 
```{r scatter-rgb-sig}
# temperature
scatter_rgb_temp <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Temperature (ºC)`,
             y = rgb_weekly, 
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (rgb_weekly - sd(rgb_weekly, na.rm = TRUE)),
                    ymax = (rgb_weekly + sd(rgb_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Pigmentation (RGB) vs Mean Temperature (ºC) Per Trial") +
  ylab("Pigmentation (RGB)") +
  xlab("Temperature (ºC)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_rgb_temp

# salinity 
scatter_rgb_sal <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Salinity (PSU)`,
             y = rgb_weekly, 
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (rgb_weekly - sd(rgb_weekly, na.rm = TRUE)),
                    ymax = (rgb_weekly + sd(rgb_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Pigmentation (RGB) vs Mean Salinity (PSU) Per Trial") +
  ylab("Pigmentation (RGB)") +
  xlab("Salinity (PSU)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_rgb_sal


# pH
scatter_rgb_pH <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `pH`,
             y = rgb_weekly, 
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (rgb_weekly - sd(rgb_weekly, na.rm = TRUE)),
                    ymax = (rgb_weekly + sd(rgb_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Pigmentation (RGB) vs Mean pH Per Trial") +
  ylab("Pigmentation (RGB)") +
  xlab("pH") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_rgb_pH


# silica
scatter_rgb_silica <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Silica`,
             y = rgb_weekly, 
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (rgb_weekly - sd(rgb_weekly, na.rm = TRUE)),
                    ymax = (rgb_weekly + sd(rgb_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Pigmentation (RGB) vs Mean Silica (mg/L) Per Trial") +
  ylab("Pigmentation (RGB)") +
  xlab("Silica (mg/L)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_rgb_silica

# iron
scatter_rgb_iron <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Iron`,
             y = rgb_weekly, 
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (rgb_weekly - sd(rgb_weekly, na.rm = TRUE)),
                    ymax = (rgb_weekly + sd(rgb_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Pigmentation (RGB) vs Mean Iron (mg/L) Per Trial") +
  ylab("Pigmentation (RGB)") +
  xlab("Iron (mg/L)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_rgb_iron

# calcium
scatter_rgb_calc <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Calcium`,
             y = rgb_weekly, 
             colour = Trial)) +
  geom_point(size = 6) +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  geom_errorbar(aes(ymin = (rgb_weekly - sd(rgb_weekly, na.rm = TRUE)),
                    ymax = (rgb_weekly + sd(rgb_weekly, na.rm = TRUE)))) +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Pigmentation (RGB) vs Mean Calcium (mg/L) Per Trial") +
  ylab("Pigmentation (RGB)") +
  xlab("Calcium (mg/L)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
scatter_rgb_calc


```

### Box Plots
#### Significant 
```{r boxplots-rgb-sig}
# temperature 
box_rgb_temp <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Temperature (ºC)`,
             y = rgb_weekly, 
             color = Trial)) +
  geom_boxplot() +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Pigmentation (RGB) vs Mean Temperature (ºC) Per Trial") +
  ylab("Pigmentation (RGB)") +
  xlab("Temperature (ºC)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
box_rgb_temp

# salinity 
box_rgb_sal <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Salinity (PSU)`,
             y = rgb_weekly, 
             color = Trial)) +
  geom_boxplot() +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Pigmentation (RGB) vs Mean Salinity (PSU) Per Trial") +
  ylab("Pigmentation (RGB)") +
  xlab("Salinity (PSU)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
box_rgb_sal

# pH 
box_rgb_pH <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `pH`,
             y = rgb_weekly, 
             color = Trial)) +
  geom_boxplot() +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Pigmentation (RGB) vs Mean pH Per Trial") +
  ylab("Pigmentation (RGB)") +
  xlab("pH") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
box_rgb_pH

# silica
box_rgb_silica <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Silica`,
             y = rgb_weekly, 
             color = Trial)) +
  geom_boxplot() +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Pigmentation (RGB) vs Mean Silica (mg/L) Per Trial") +
  ylab("Pigmentation (RGB)") +
  xlab("Silica (mg/L)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
box_rgb_silica

# iron
box_rgb_iron <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Iron`,
             y = rgb_weekly, 
             color = Trial)) +
  geom_boxplot() +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Pigmentation (RGB) vs Mean Iron (mg/L) Per Trial") +
  ylab("Pigmentation (RGB)") +
  xlab("Iron (mg/L)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
box_rgb_iron

# calcium
box_rgb_calc <- dfCoralRateAllEnvMean %>%
  ggplot(aes(x = `Calcium`,
             y = rgb_weekly, 
             color = Trial)) +
  geom_boxplot() +
  geom_smooth(method = "lm", se = FALSE, col = 'red') +
  scale_color_manual(values = c("cornflowerblue", "dodgerblue", "deepskyblue2", "darkorange3", "azure3", "springgreen3", "purple")) +
  ggtitle("Pigmentation (RGB) vs Mean Calcium (mg/L) Per Trial") +
  ylab("Pigmentation (RGB)") +
  xlab("Calcium (mg/L)") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
box_rgb_calcium

```


# References
weighted least squares regression: <https://www.statology.org/weighted-least-squares-in-r/>