---
title: "Mini Ocean Lighting Project Stats: PERMANOVA"
collaborators: Savanna Saunders, Renee Grambihler
project: Mini Ocean Light Data
institution: Biosphere 2 Ocean
description: "Uses non parametric statistics to analyze size, weight, and color changes or corals grown in Mini Ocean under different lighting treatments from Summer 2024 to Spring 2025 Coral Lighting Experiment"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

Input: 
  `CoralColorWeightData_Rates.csv`: made as output in `MO-Wrangling-PERMANOVA.Rmd`, coral weight, surface area, and rgb (color) data as weekly rate of change per trial; one data point per coral per trial indicating the total change in response variable per trial
  
Output: 
  `MOStats_PERMANOVA_SingleIndependent.xlsx`: results of running a single PERMANOVA with no interactions between the independent variables. 
  
  `MOStats_PERMANOVA_SingleInteraction.xlsx`: results of running a single PERMANOVA with interactions between the independent variables. 
  
  `MOStats_PERMANOVA_Pairwise_All.xlsx`: results of Pairwise PERMANOVA tests run on Trial, Treatment, and Species separately. 
  
  `MOStats_TukeyTestResults.xlsx`: results of Tukey Test run with interactions between the independent variables. 
  

## Import Libraries
```{r import-libraries}
library(tidyverse)
library(readxl)
library(openxlsx)
library(rstatix)
library(vegan)
library(multcompView)
library(pairwiseAdonis) 
# to install pairwiseAdonis: in console run install.packages('devtools') > library(devtools) > install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
options(scipen = 9999)
# Do this to avoid a stupid useless error message
options(dplyr.summarise.inform = FALSE)  
```

## Import Data
```{r import-data}
df_rates <- read_csv("InputData/Clean/CoralData/CoralColorWeightData_Rates.csv") %>%
  # get rid of trial PBR 4 due to light timer malfunction keeping lights on continuously for 10 days
  filter(Trial != "PBR 4")

```

# Quick Data Distribution Visualization
## Trial
```{r distribution-rate-data-histogram-trial}
# Weight
df_rates %>%
  ggplot(aes(x = weight_g_weekly,
             fill = Treatment,
             color = Treatment)) +
  facet_grid(rows = vars(Trial)) +
  geom_histogram() +
  # add dotted line at x = 0
  geom_vline(xintercept = 0, linetype = "dotted") +
  ggtitle("Weekly Weight (g)") +
  scale_fill_manual(values = c("cyan3", "red2", "grey")) +
  scale_color_manual(values = c("cyan4", "red3", "darkgrey")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())

# Surface Area
df_rates %>%
  ggplot(aes(x = surface_area_cm3_weekly,
             fill = Treatment,
             color = Treatment)) +
  facet_grid(rows = vars(Trial)) +
  geom_histogram() +
  # add dotted line at x = 0
  geom_vline(xintercept = 0, linetype = "dotted") +
  ggtitle("Surface Area (cm^3)") +
  scale_fill_manual(values = c("cyan3", "red2", "grey")) +
  scale_color_manual(values = c("cyan4", "red3", "darkgrey")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())

# RGB
df_rates %>%
  ggplot(aes(x = rgb_weekly,
             fill = Treatment,
             color = Treatment)) +
  facet_grid(rows = vars(Trial)) +
  geom_histogram() +
  # add dotted line at x = 0
  geom_vline(xintercept = 0, linetype = "dotted") +
  ggtitle("Weekly Pigmentation (RGB Values)") +
  scale_fill_manual(values = c("cyan3", "red2", "grey")) +
  scale_color_manual(values = c("cyan4", "red3", "darkgrey")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())

```

## Species and Trial
```{r distribution-rate-data-histogram-species-trial}
# Weight
df_rates %>%
  ggplot(aes(x = weight_g_weekly,
             fill = Treatment,
             color = Treatment)) +
  facet_grid(rows = vars(Trial),
             cols = vars(Species)) +
  geom_histogram() +
  # add dotted line at x = 0
  geom_vline(xintercept = 0, linetype = "dotted") +
  ggtitle("Weekly Weight (g)") +
  scale_fill_manual(values = c("cyan3", "red2", "grey")) +
  scale_color_manual(values = c("cyan4", "red3", "darkgrey")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())

# surface area
df_rates %>%
  ggplot(aes(x = surface_area_cm3_weekly,
             fill = Treatment,
             color = Treatment)) +
  facet_grid(rows = vars(Trial),
             cols = vars(Species)) +
  geom_histogram() +
  # add dotted line at x = 0
  geom_vline(xintercept = 0, linetype = "dotted") +
  ggtitle("Weekly Weight (g)") +
  scale_fill_manual(values = c("cyan3", "red2", "grey")) +
  scale_color_manual(values = c("cyan4", "red3", "darkgrey")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())

# rgb
df_rates %>%
  ggplot(aes(x = rgb_weekly,
             fill = Treatment,
             color = Treatment)) +
  facet_grid(rows = vars(Trial),
             cols = vars(Species)) +
  geom_histogram() +
  # add dotted line at x = 0
  geom_vline(xintercept = 0, linetype = "dotted") +
  ggtitle("Weekly Weight (g)") +
  scale_fill_manual(values = c("cyan3", "red2", "grey")) +
  scale_color_manual(values = c("cyan4", "red3", "darkgrey")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())


```

# PERMANOVAs
Run non-parametric multivariate analysis of variance (NPMANOVA), also known as permutational multivariate analysis of variance (PERMANOVA), using weight, surface area, and RGB as dependent variables and Treatment and Species as independent/grouping/factor variables. The goal here is to see if there is a significant difference in the variance of the dependent variables as grouped by the independent variables. First we'll test Treatment, then Species, then Treatment and Species. 

Note: we are using euclidean distance instead of curtis-bray distance due to having negative values in our dependent data. 

We will also need to set the seed for the number generator used within the PERMANOVA function to ensure the same sequence of pseudo-random permutations is used each time we run the code, making our results reproducible - the specific number we choose for the seed is arbitrary. We will need to do this each time before we run a PERMANOVA.

On the F Statistic: "When the F value is large, it indicates that the differences in group means are substantially greater than what could be expected due to random variation alone. Conversely, a small F value suggests that the observed differences in means are likely attributable to chance fluctuations within the groups." (<https://www.6sigma.us/six-sigma-in-focus/anova-f-value-meaning/>)

## Single PERMANOVAs
### Set-up
Make a dataframe holding just the dependent variables.
```{r permanova-setup}
# from <https://f-santos.gitlab.io/2020-05-07-npmanova.html>
y <- df_rates[, c("weight_g_weekly", "surface_area_cm3_weekly", "rgb_weekly")]

```

### Test Terms Independently 
Test the effects of each factor variable on the dependent data independently of each other. 
```{r permanova-independent-terms}
# set seed for permutations so that they are reproducible 
set.seed(100)

# regress y on Trial, Species, Treatment
perm_independent <- adonis2(y ~ df_rates$Species + df_rates$Treatment + df_rates$Trial, 
                            method = "euclidean")

# add column with row names that will print out nicely
df_perm_independent <- perm_independent %>%
  mutate(ModelNames = rownames(perm_independent), .before = Df)

# make excel file of results
#write.xlsx(x = df_perm_independent, file = "OutputData/StatisticsResults/MOStats_PERMANOVA_SingleIndependent.xlsx")

```
Species: 
  - p = 0.002
  - F = 5.964463
  - R^2 = 0.04125117

Treatment: 
  - p = 0.001
  - F = 8.024739
  - R^2 = 0.05550036
  
Trial: 
  - p = 0.001
  - F = 11.549877
  - R^2 = 0.15976153
  
Residual R^2 = 0.74348694

There is a significant difference in dependent variable variance based on Treatment, Species, and Trial independently of each other. Run `view(perm_independent)` to see full table of results. 

### Test Interaction Between Terms
```{r permanova-interaction-terms}
# set seed for permutations so that they are reproducible 
set.seed(100)

# regress y on Trial, Species, Treatment, Trial*Species, Trial*Treatment, Species*Treatment
perm_interaction <- adonis2(y ~ df_rates$Species * df_rates$Treatment * df_rates$Trial, 
                            method = "euclidean")

# add column with row names that will print out nicely
df_perm_interaction <- perm_interaction %>%
  mutate(ModelNames = rownames(perm_interaction), .before = Df)

# make excel file of results
#write.xlsx(x = df_perm_interaction, file = "OutputData/StatisticsResults/MOStats_PERMANOVA_SingleInteraction.xlsx")

```
Species: 
  - p = 0.001
  - F = 7.082575
  - R^2 = 0.04125117

Treatment: 
  - p = 0.001
  - F = 9.529075
  - R^2 = 0.05550036

Trial: 
  - p = 0.001
  - F = 13.715042
  - R^2 = 0.15976153

Species & Treatment: 
  - p = 0.026
  - F = 2.763092
  - R^2 = 0.03218625

Species & Trial: 
  - p = 0.001
  - F = 5.156502
  - R^2 = 0.12013243
  
Residual R^2 = 0.59116826

There is a significant difference in dependent variable variance based on Treatment, Species, Trial, Treatment * Species, and Species * Trial. 

NOTE: Treatment*Trial was automatically removed due to high multicollinearity, meaning that the variables are very highly correlated, which makes it difficult for the model to discern their individual impacts. Run `view(perm_interaction)` to see full table of results.

## Pairwise PERMANOVAs
Next we'll perform pairwise PERMANOVAs to see if there is a significant difference in the variance of the dependent variables as grouped by the specific levels of the independent variables. First we'll test Treatment, then Species, then Treatment and Species.

### Setup
```{r pairwise-permanova-setup-1}
# R sometimes has trouble with assigning proper formats to variables so this sets it as a factor (the proper format)
df_rates$Species <- as.factor(df_rates$Species)
df_rates$Treatment <- as.factor(df_rates$Treatment)
df_rates$Trial <- as.factor(df_rates$Trial)

# then we'll use the same y dataframe from the regular permanova step
y <- df_rates[, c("weight_g_weekly", "surface_area_cm3_weekly", "rgb_weekly")]
```

### Test Terms Independently
#### Species
```{r pairwise-permanova-independent-species}
# set seed for permutations so that they are reproducible 
set.seed(100)

# Species model
perm_pair_species <- pairwise.adonis2(y ~ Species, data = df_rates, method = "euclidean")

# View Mcav vs Ofav
perm_pair_species$Mcav_vs_Ofav
# View Mcav vs Pcli
perm_pair_species$Mcav_vs_Pcli
# View Ofav vs Pcli
perm_pair_species$Ofav_vs_Pcli

# put results into dataframe 
# list of comparisons
comparisons_species <- names(perm_pair_species)[2:4] #element 1 is the parent call which we don't need
# empty dataframe to put results in
df_PermPairSpecies <- data.frame()


for (i in 1:length(comparisons_species)) {
  # set counter as i + 1 in order to avoid grabbing parent call info from perm_pair_trial (element)
  counter <- i + 1
  
  # set comparison
  comparison <- comparisons_species[i]
  
  # grab first row of comparison model stats
  stats <- perm_pair_species[[counter]][1,]
  
  # combine comparison and stats into dataframe
  perm <- cbind(comparison, stats)
  
  # add perm dataframe to df_PermPairTrial dataframe 
  df_PermPairSpecies <- rbind(df_PermPairSpecies, perm)
}

# remove row names
rownames(df_PermPairSpecies) <- NULL
# rename p-value column
df_PermPairSpecies <- rename(df_PermPairSpecies, p = `Pr(>F)`)
# round numeric columns to 4 digits
df_PermPairSpecies <- df_PermPairSpecies %>%
  mutate(across(where(is.numeric), round, digits = 4)) %>%
  add_column(term = "Species", .before = "comparison")

```
Mcav vs Ofav: 
  - p = 0.002
  - F = 7.8457
  - R^2 = 0.05034
  
Mcav vs Pcli: 
  - p = 0.578
  - F = 0.3032
  - R^2 = 0.00206
  
Ofav vs Pcli: 
  - p = 0.007
  - F = 6.759
  - R^2 = 0.04396

#### Treatment
```{r pairwise-permanova-independent-treatment}
# set seed for permutations so that they are reproducible 
set.seed(100)

# Treatment model 
perm_pair_treatment <- pairwise.adonis2(y ~ Treatment, data = df_rates, method = "euclidean")

# View PBR vs Control
perm_pair_treatment$PBR_vs_Control
# View Control vs PBC
perm_pair_treatment$Control_vs_PBC
# View PBR vs PBC
perm_pair_treatment$PBR_vs_PBC

# put results into dataframe 
# list of comparisons
comparisons_treatment <- names(perm_pair_treatment)[2:4] #element 1 is the parent call which we don't need
# empty dataframe to put results in
df_PermPairTreatment <- data.frame()


for (i in 1:length(comparisons_treatment)) {
  # set counter as i + 1 in order to avoid grabbing parent call info from perm_pair_trial (element)
  counter <- i + 1
  
  # set comparison
  comparison <- comparisons_treatment[i]
  
  # grab first row of comparison model stats
  stats <- perm_pair_treatment[[counter]][1,]
  
  # combine comparison and stats into dataframe
  perm <- cbind(comparison, stats)
  
  # add perm dataframe to df_PermPairTrial dataframe 
  df_PermPairTreatment <- rbind(df_PermPairTreatment, perm)
}

# remove row names
rownames(df_PermPairTreatment) <- NULL
# rename p-value column
df_PermPairTreatment <- rename(df_PermPairTreatment, p = `Pr(>F)`)
# round numeric columns to 4 digits
df_PermPairTreatment <- df_PermPairTreatment %>%
  mutate(across(where(is.numeric), round, digits = 4))%>%
  add_column(term = "Treatment", .before = "comparison")

```
PBR vs Control: 
  - p = 0.01
  - F = 7.483
  - R^2 = 0.0504
  
PBC vs Control: 
  - p = 0.003
  - F = 11.09
  - R^2 = 0.07244
  
PBR vs PBC: 
  - p = 0.302
  - F = 1.0887
  - R^2 = 0.0068

  
#### Trial
```{r pairwise-permanova-independent-trial}
# set seed for permutations so that they are reproducible 
set.seed(100)

# Trial model 
perm_pair_trial <- pairwise.adonis2(y ~ Trial, data = df_rates, method = "euclidean")

# put results into dataframe 
# list of comparisons
comparisons_trial <- names(perm_pair_trial)[2:22] #element 1 is the parent call which we don't need
# empty dataframe to put results in
df_PermPairTrial <- data.frame()


for (i in 1:length(comparisons_trial)) {
  # set counter as i + 1 in order to avoid grabbing parent call info from perm_pair_trial (element)
  counter <- i + 1
  
  # set comparison
  comparison <- comparisons_trial[i]
  
  # grab first row of comparison model stats
  stats <- perm_pair_trial[[counter]][1,]
  
  # combine comparison and stats into dataframe
  perm <- cbind(comparison, stats)
  
  # add perm dataframe to df_PermPairTrial dataframe 
  df_PermPairTrial <- rbind(df_PermPairTrial, perm)
}

# remove row names
rownames(df_PermPairTrial) <- NULL
# rename p-value column
df_PermPairTrial <- rename(df_PermPairTrial, p = `Pr(>F)`)
# round numeric columns to 4 digits
df_PermPairTrial <- df_PermPairTrial %>%
  mutate(across(where(is.numeric), round, digits = 4))%>%
  add_column(term = "Trial", .before = "comparison")

```
Significant Results: 
  *Control vs PBC*: 
    Control_vs_PBC 3
      - p = 0.001
      - F = 39.8600
      - R^2 = *0.3118*
    
    Control_vs_PBC 1
      - p = 0.008
      - F = 7.0570
      - R^2 = 0.0742
  
  *Control vs PBR*: 
    Control_vs_PBR 2
      - p = 0.006
      - F = 7.0970
      - R^2 = 0.0754
      
  *PBR vs PBC*: 
    PBR 1_vs_PBC 3
      - p = 0.001
      - F = 25.5700
      - R^2 = *0.3296*
      
    PBR 3_vs_PBC 3
      - p = 0.001
      - F = 18.1600
      - R^2 = 0.2588  
      
    PBC 2_vs_PBR 2
      - p = 0.003
      - F = 10.2400
      - R^2 = 0.1672
      
    PBC 2_vs_PBR 2
      - p = 0.003
      - F = 10.2400
      - R^2 = 0.1672
      
    PBR 2_vs_PBC 3
      - p = 0.009
      - F = 8.8360
      - R^2 = 0.1477
      
    PBR 1_vs_PBC 2
      - p = 0.017
      - F = 6.6390
      - R^2 = 0.1132
    
    PBC 2_vs_PBR 3
      - p = 0.025
      - F = 5.1780
      - R^2 = 0.0906
    
  *PBR vs PBR*: 
    None
  
  *PBC vs PBC*: 
    *PBC 2_vs_PBC 3*
      - p = 0.001
      - F = 44.2900
      - R^2 = *0.4600*
      
    PBC 1_vs_PBC 3
      - p = 0.002
      - F = 11.4000
      - R^2 = 0.1799
    
    PBC 1_vs_PBC 2
      - p = 0.003
      - F = 11.2000
      - R^2 = 0.1772

#### All Pairwise PERMANOVA Results 
```{r pairwise-permanova-results-all}
# combine all pairwise permanova results into one data frame
dfPairPermAll <- rbind(df_PermPairSpecies, df_PermPairTreatment, df_PermPairTrial)

# write excel file of results
# write.xlsx(x = dfPairPermAll, file = "OutputData/StatisticsResults/MOStats_PERMANOVA_Pairwise_All.xlsx")

```

# Post Hoc Tests: Tukey
For each of the 3 response variables, we are going to run Tukey's Honest Significant Difference (Tukey HSD) test on all 3 independent variables and their significant interactions from the single PERMANOVAs: Species, Trial, Treatment, Species:Treatment, and Species:Trial. 

## Data Set-Up
```{r post-hoc-tukey-data-setup}
# set up data frame for tukey testing 
df_rates_long <- df_rates %>%
  pivot_longer(cols = 5:7,
               names_to = "dependent_variable_names",
               values_to = "dependent_variable_values")

```

## Run Tukey Test
```{r tukey-test-treatment-species}
tukey_treat_tri_spec <- df_rates_long %>%
  group_by(dependent_variable_names) %>%
  tukey_hsd(dependent_variable_values ~ Treatment*Species*Trial) %>%
  select(-null.value) %>%
  mutate(across(where(is.numeric), round, digits = 4)) 

# view results
tukey_treat_tri_spec

# make an excel file of results
# write.xlsx(x = tukey_treat_tri_spec, file = "OutputData/StatisticsResults/MOStats_TukeyTestResults.xlsx")

```

## Degrees of Freedom
### Weight
```{r post-hoc-tukey-extra-stats}
summary(aov(weight_g_weekly ~ Treatment * Species * Trial, data = df_rates))
```
The degrees of freedom for a Tukey HSD test are the df of the Residuals from the ANOVA test that the Tukey test is based on, so 203 in this case. 

# Plots
```{r plots}
#### pbc by trial #### 
# pbc 1 = june/july
# pbc 2 = nov/dec
# pbc 3 = feb/march

# rgb 
df_rates %>%
  filter(Treatment == "PBC") %>%
  ggplot(aes(y = rgb_weekly, 
             color = Trial)) +
  geom_boxplot()

# weight 
df_rates %>%
  filter(Treatment == "PBC") %>%
  ggplot(aes(y = weight_g_weekly, 
             color = Trial)) +
  geom_boxplot()

# surface area  
df_rates %>%
  filter(Treatment == "PBC") %>%
  ggplot(aes(y = surface_area_cm3_weekly, 
             color = Trial)) +
  geom_boxplot()

# dfMultManovaTest_long_rgb <- dfMultManovaTest %>%
#   pivot_longer(cols = rgb_mean:rgb_mean,
#                names_to = "dependent_variables",
#                values_to = "values")
#   
#   dfMultManovaTest_long_rgb %>%
#   ggplot(aes(x = time,
#              y = values, 
#              fill = dependent_variables)) +
#   geom_boxplot() +
#   facet_wrap(~treatment) +
#   theme(panel.grid = element_blank())
#   
#   
#   par_summaryall <- parFrame_all %>%
#   group_by(Treatment) %>%
#   mutate(Treatment = factor(Treatment, levels = c("PBR-1", "PBC-1", "PBR-2", "PBC-2", "PBR-3", "PBC-3", "PBR-4"))) %>%
#   summarise(mean_par = mean(PAR, na.rm = TRUE),
#             sd_par = sd(PAR, na.rm = TRUE), #standard deviation
#             n = n(),
#             se_par = sd_par / sqrt(n)) #standard error
# 
# ggplot(par_summaryall, aes(x = Treatment, y = mean_par, fill = Treatment)) +
#   geom_boxplot() +
#   geom_errorbar(aes(ymin = mean_par - se_par, ymax = mean_par + se_par), width = 0.2) +
#   ylim(0, NA)+
#   labs(title = "Mean PAR by Treatment",
#        x = "Treatment",
#        y = "Mean PAR") +
#   theme_minimal()
```


# References
PERMANOVAs: 
 - <https://f-santos.gitlab.io/2020-05-07-npmanova.html>
 - <https://www.6sigma.us/six-sigma-in-focus/anova-f-value-meaning/>
Formulas used in PERMANOVAs: 
  - <https://www.econometrics.blog/post/the-r-formula-cheatsheet/>
Tukey Test: 
  - Abdi, H. and Williams, L.J. (2010) Turkey’s Honestly Significant Difference (HSD) test. Encyclopedia of Research Design. Sage, Thousand Oaks, 1-5.
