---
title: "Mini Ocean Lighting Project: Figures"
author: Savanna Saunders, Renee Grambihler
project: Mini Ocean Light Data
institution: Biosphere 2 Ocean
description: "Figures created from non parametric statistical analysis and raw data to analyze size, weight, and color changes or corals grown in Mini Ocean under different lighting treatments for Summer 2024 to Spring 2025 Coral Lighting Experiment"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

**Input**:
  `CoralColorWeightData_Rates.csv`: coral data with Before and After values subtracted from each other and then divided either by 3 or 6 depending on what Trial the data is from in order to get a weekly rate of change. 

  `All-Environmental-Variables-Wide.xlsx`: all environmental data (nutrients, water quality, DLI) from MO and R2 during all trials in wide format

  `All-Environmental-Variables-Long.xlsx`: all environmental data (nutrients, water quality, DLI) from MO and R2 during all trials in long format

  `MOStats_PERMANOVA_Pairwise_All.xlsx`: results of Pairwise PERMANOVA tests run on Trial, Treatment, and Species separately. 

  `MOStats_TukeyTestResults.xlsx`: results of Tukey Test run with interactions between the independent variables.

**Output**:
  `HeatMap-PairwisePERMANOVA.png`: heat map of pairwise PERMANOVA results by treatment, trial, and species.
  
  `HeatMap-TukeyTest-Treatment.png`: heat map of tukey test results by treatment.
  
  `HeatMap-TukeyTest-TreatmentSpecies-Between.png`: heat map of tukey test results between groups factors for treatment and species.
  
  `HeatMap-TukeyTest-TreatmentSpecies-Within.png`: heat map of tukey test results within groups factors for treatment and species. 
  
  `BoxPlot-All-Treatment-Species.png`: box plot of all response variables by treatment and species. 
  
  `BoxPlot-DLI-Per-Trial.png`: box plot of daily light integral (DLI) values by trial.
  
  `BoxPlot-RGB-PBCTrials.png`: box plot of RGB values for the PBC trials.
  
  `ConfoundingFactors-Ranges-Stated.xlsx`: table of ranges for confounding factor variables stated as "minimum - maximum"

  `ConfoundingFactors-Ranges-Difference.xlsx`: table of ranges for confounding factor varibles as the difference between the maximum and minimum values. 



```{r import-libraries}
library(tidyverse)
library(readxl)
library(rstatix)
library(lubridate)
library(lmtest)
library(modelr)
library(ggpubr)
library(radiant.data)
# get rid of scientific notation
options(scipen = 9999)
# avoid dplyr error message
options(dplyr.summarise.inform = FALSE)  
```

# Import Data and Wrangle
Data wrangling includes adding columns for labeling and other figure aesthetics, but most important is **multiplying RGB values by -1** so that decreases in RGB correspond to decreases in pigmentation and figures will not have to be read inversely. 
```{r new-data-import}
# coral data 
dfCoral <- read_csv("InputData/Clean/CoralData/CoralColorWeightData_Rates.csv") %>%
  # get rid of trial PBR 4 due to light timer malfunction keeping lights on continuously for 10 days
  filter(Trial != "PBR 4") %>%
  #make species names paper-ready
  mutate(Species = case_when(
    Species == "Mcav" ~ "M. cavernosa",
    Species == "Ofav" ~ "O. faveolata",
    Species == "Pcli" ~ "P. clivosa"
  )) %>%
  # change pigmentation signs to make reading the figure more intuitive since negative values indicated increased pigmentation 
  mutate(rgb_weekly = rgb_weekly * -1)

# environmental data
# wide / regular 
dfEnvVar <- read_excel("InputData/Clean/EnvironmentalData/All-Environmental-Variables-Wide.xlsx")
# long
dfEnvVar_long <- read_excel("InputData/Clean/EnvironmentalData/All-Environmental-Variables-Long.xlsx")

# results files
# pairwise permanova 
dfPairPerm <- read_excel("OutputData/StatisticsResults/MOStats_PERMANOVA_Pairwise_All.xlsx") %>%
  #substitute _ for white space in comparison column for plotting aesthetics 
  mutate(comparison = gsub("_", " ", comparison),
         #change order of some comparison observations for plotting aesthetics 
         comparison = gsub("PBR vs Control", "Control vs PBR", comparison),
         comparison = gsub("PBR 1 vs Control", "Control vs PBR 1", comparison),
         comparison = gsub("PBC 1 vs PBR 2", "PBR 2 vs PBC 1", comparison),
         comparison = gsub("PBC 1 vs PBR 3", "PBR 3 vs PBC 1", comparison),
         comparison = gsub("PBC 2 vs PBR 2", "PBR 2 vs PBC 2", comparison),
         comparison = gsub("PBC 2 vs PBR 3", "PBR 3 vs PBC 2", comparison)) %>%
  #change species names to look correct on plots 
  mutate(comparison = gsub("Mcav", "M. cavernosa", comparison),
         comparison = gsub("Ofav", "O. faveolata", comparison),
         comparison = gsub("Pcli", "P. clivosa", comparison)) %>%
  mutate(term = factor(term, levels = c("Treatment", "Trial", "Species"))) %>%
  #add column with significance stars 
  mutate(p_sig = case_when(
    p < 0.001 ~ "***",
    p >= 0.001 & p < 0.01 ~ "**",
    p >= 0.01 & p < 0.05 ~ "*",
    p >= 0.05 ~ "", 
  ))

# tukey 
dfTukey <- read_excel("OutputData/StatisticsResults/MOStats_TukeyTestResults.xlsx") %>%
  #change dependent variable names to look better on plots 
  mutate(dependent_variable_names = case_when(
    dependent_variable_names == "rgb_weekly" ~ "Pigmentation (RGB)",
    dependent_variable_names == "surface_area_cm3_weekly" ~ "Surface Area (cm^3)",
    dependent_variable_names == "weight_g_weekly" ~ "Weight (g)"
  )) %>%
  # change pigmentation signs to make reading the figure more intuitive since negative values indicated increased pigmentation 
  mutate(estimate = ifelse(dependent_variable_names == "Pigmentation (RGB)",
                           estimate * -1,
                           estimate))


```

# Color Palettes
Set up color palettes we might want to use. There's cool, warm, and a few specific ones from the Virids package. All of these are colorblind friendly. 
```{r color-options}
#ColorBrewer 
# warm palette
cols_warm <- c("#feedde", "#fdbe85", "#fd8d3c", "#e6550d", "#a63603")

# cool palette
cols_cool <- c("#f1eef6", "#d7b5d8", "#df65b0", "#dd1c77", "#980043")

# Viridis
# rocket palette full
cols_rocket <- c("#FAEBDDFF","#F4875EFF","#CB1B4FFF","#611F53FF","#03051AFF")

# rocket palette 3 colors
cols_rocket_sm <- c("#611F53FF","#F4875EFF","#FAEBDDFF")

# rocket pallette 6 diverging colors 
cols_rocket_div <- c("#a63603","#F4875EFF","#FAEBDDFF","#CB1B4FFF","#611F53FF","#03051AFF")

```


# Heat Maps
Similar to figure found in Hackerott et al. 2025: <https://doi.org/10.21203/rs.3.rs-5247850/v1>
## Pairwise PERMANOVA
Plot pairwise PERMANOVA results by treatment, trial, and species. Significant results are indicated by the presence of stars on each cell. Darker colors indicate a higher r^2 value. 
```{r heat-map-pairwise-permanova}
# order comparison column by term column factors so it looks nice in the plot
dfPairPerm$comparison <- factor(dfPairPerm$comparison, levels = (dfPairPerm$comparison)[order(dfPairPerm$term, decreasing = TRUE)])

heatmap_PairPerm <- dfPairPerm %>%
  ggplot(aes(x = term,
             y = comparison, 
             fill = R2)) +
  geom_tile(color = "white",
            lwd = 0.5,
            linetype = 1) +
  geom_text(aes(label = p_sig), color = "white") +
  labs(fill = expression(r^2)) +
  scale_fill_stepsn(colors = cols_rocket) +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        panel.border = element_blank(),
        axis.line.x.bottom = element_line(),
        axis.line.y.left = element_line(),
        axis.text.x = element_text(angle = 45, hjust = 1, color = "black"),
        axis.text.y = element_text(color = "black"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

heatmap_PairPerm

# save plot 
# ggsave("Figures/StatisticsResults/HeatMaps/HeatMap-PairwisePERMANOVA.png", plot = heatmap_PairPerm)

```

## Tukey Test
### Treatment
Plot tukey test results by Treatment. Y-axis shows Treatment comparisons as Treatment 1 minus (vs) Treatment 2, and x-axis shows each of the dependent variables. Darker colors indicate a higher/positive difference, and lighter colors indicate a lower/negative difference between treatments. For example, PBC was more pigmented than Control. 
#### Data Preparation 
```{r heat-map-tukey-treatment-df}
# data wrangling
dfHeatMap_Tukey_Treat <- dfTukey %>%
  filter(term == "Treatment") %>%
  # estimate comes from group 2 - group 1
  mutate(comparison = paste(group2, "vs", group1, sep = " ")) %>%
  # remove unused columns
  select(-group2, -group1, -conf.low, -conf.high) %>% 
  # remove 'ns' to clean up the plot aesthetically  
  mutate(p.adj.signif = ifelse(p.adj.signif == "ns", "", p.adj.signif)) %>%
  # normalize estimates to same scale so we can see the differences in weight and surface area rather than just pigment since the RGB values are much larger than the weight and surface area values  
  # normalize to [-1,1] 
  mutate(z_score = (estimate - mean(estimate)) / sd(estimate), 
         min_max = (2 * (estimate - min(estimate)) / 
                           (max(estimate) - min(estimate)) -1), 
         .after = estimate)

```

#### Plot 
```{r heat-map-tukey-treat-plot}
# plot data 
heatmap_tukey_treat <- dfHeatMap_Tukey_Treat %>%
  ggplot(aes(x = dependent_variable_names,
             y = comparison, 
             fill = estimate)) +
  geom_tile(color = "white",
            lwd = 0.5,
            linetype = 1) +
  coord_fixed() +
  # put estimate values and significance stars on grid boxes 
  geom_text(aes(label = paste(p.adj.signif,
                              round(estimate, digits = 4), sep = " ")),
            color = "white",
            size = 3) +
  scale_x_discrete(labels = c("Pigmentation (RGB)", expression(paste("Surface Area (c", m^3, ")", sep = "")), "Weight (g)")) +
  scale_fill_stepsn(
    name = "Estimate",
    colors = cols_rocket
  ) +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_rect(fill = "white"), 
        axis.line = element_blank(),
        axis.text.x = element_text(angle = 45, 
                                   hjust = 1, 
                                   color = "black", 
                                   size = 8),
        axis.text.y = element_text(color = "black", size = 8),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 9))

# view plot
heatmap_tukey_treat

#save plot 
# ggsave("Figures/StatisticsResults/HeatMaps/HeatMap-TukeyTest-Treatment.png",
#        plot = heatmap_tukey_treat,
#        height = 4, width = 4)
```


### Treatment and Species: Between Groups
Create a heatmap to plot between groups differences. The values shown in each cell are the differences between the response variables for comparisons of the 3 treatments, with each panel representing one Species. For example, the first row shows the difference in estimate between PBR and PBC for each of the response variables, which is then faceted by Species, so the first panel is for Mcav.
#### Data Preparation
```{r heat-map-tukey-treatment-species-btwn-df}
dfTukey_TreatSpec <- dfTukey %>%
  filter(term == "Treatment:Species") %>%
  separate(col = group1, into = c("group1.treatment", "group1.species"), sep = ":") %>%
  separate(col = group2, into = c("group2.treatment", "group2.species"), sep = ":") %>%
  mutate(group1.species = ifelse(group1.species == "Mcav", "M. cavernosa", group1.species),
        group1.species = ifelse(group1.species == "Ofav", "O. faveolata", group1.species),
        group1.species = ifelse(group1.species == "Pcli", "P. clivosa", group1.species)) %>%
  mutate(group2.species = ifelse(group2.species == "Mcav", "M. cavernosa", group2.species),
        group2.species = ifelse(group2.species == "Ofav", "O. faveolata", group2.species),
        group2.species = ifelse(group2.species == "Pcli", "P. clivosa", group2.species)) 
  
# subset to compare same species in different treatments 
# mcav 
dfTukey_mcav <- dfTukey_TreatSpec %>%
  filter(group1.species == "M. cavernosa" & group2.species == "M. cavernosa")
# ofav
dfTukey_ofav <- dfTukey_TreatSpec %>%
  filter(group1.species == "O. faveolata" & group2.species == "O. faveolata")
# pcli 
dfTukey_pcli <- dfTukey_TreatSpec %>%
  filter(group1.species == "P. clivosa" & group2.species == "P. clivosa")

# combine all direct species comparisons 
dfTukey_TreatSpec <- rbind(dfTukey_mcav, dfTukey_ofav, dfTukey_pcli) %>%
  # use group1.species as species column to facet by in plot 
  mutate(species = group1.species) %>%
  # make comparison column of treatment 1 vs treatment 2: estimate comes from group 2-group 1
  mutate(comparison = paste(group2.treatment, "vs", group1.treatment, sep = " ")) %>%
  rename(Estimate = estimate) %>%
  mutate(p.adj.signif = ifelse(p.adj.signif == "ns", "", p.adj.signif)) 


```

#### Plot 
```{r heat-map-tukey-treatment-species-btwn-plot}
# plot
HeatMap_Tukey_TreatSpec_btwn <- dfTukey_TreatSpec %>%
  ggplot(aes(x = dependent_variable_names,
             y = comparison, 
             fill = Estimate)) +
  geom_tile(color = "white",
            lwd = 0.5,
            linetype = 1) +
  coord_fixed() +
  facet_wrap(~species) +
  geom_text(aes(label = paste(p.adj.signif, round(Estimate, digits = 3), sep = " ")), color = "white") +
  scale_x_discrete(labels = c("Pigmentation (RGB)", expression(paste("Surface Area (c", m^3, ")", sep = "")), "Weight (g)")) +
  scale_fill_stepsn(
    name = "Estimate",
    colors = cols_rocket
  ) +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line = element_blank(),
        axis.text.x = element_text(angle = 45, 
                                   hjust = 1, 
                                   color = "black",
                                   size = 9),
        axis.text.y = element_text(color = "black", size = 9),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(), 
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "italic", size = 9),
        legend.title = element_text(size = 9))
  
HeatMap_Tukey_TreatSpec_btwn

#save plot 
# ggsave("Figures/StatisticsResults/HeatMaps/HeatMap-TukeyTest-TreatmentSpecies-Between.png",
#        plot = HeatMap_Tukey_TreatSpec_btwn,
#        width = 8, height = 4)

```

### Treatment and Species: Within Groups
Create a heatmap to plot within groups differences. The values shown in each cell are the differences between the response variable values for comparisons of the 3 Species, with each panel representing one Treatment. For example, the first row shows the difference in estimate between Pcli and Ofav for each of the response variables, which is then faceted by Treatment, so the first panel is Control. 
#### Data Preparation 
```{r heat-map-tukey-treatment-species-within-df}
dfTukey_Treat_within <- dfTukey %>%
  # get just Treatment:Species
  filter(term == "Treatment:Species") %>%
  # separate group1 and group2 columns into treatment and species
  separate(col = group1, into = c("group1.treatment", "group1.species"), sep = ":") %>%
  separate(col = group2, into = c("group2.treatment", "group2.species"), sep = ":") %>%
  # change the names of the species to be better for the plots
  mutate(group1.species = ifelse(group1.species == "Mcav", "M. cavernosa", group1.species),
        group1.species = ifelse(group1.species == "Ofav", "O. faveolata", group1.species),
        group1.species = ifelse(group1.species == "Pcli", "P. clivosa", group1.species)) %>%
  mutate(group2.species = ifelse(group2.species == "Mcav", "M. cavernosa", group2.species),
        group2.species = ifelse(group2.species == "Ofav", "O. faveolata", group2.species),
        group2.species = ifelse(group2.species == "Pcli", "P. clivosa", group2.species))

# subset to compare different species in same treatments
# Control
dfTukey_within_control <- dfTukey_Treat_within %>%
  filter(group1.treatment == "Control" & group2.treatment == "Control")
# PBR
dfTukey_within_pbr <- dfTukey_Treat_within %>%
  filter(group1.treatment == "PBR" & group2.treatment == "PBR")
# PBC
dfTukey_within_pbc <- dfTukey_Treat_within %>%
  filter(group1.treatment == "PBC" & group2.treatment == "PBC")

# combine all direct treatment comparisons
dfTukey_Treat_within <- rbind(dfTukey_within_control, dfTukey_within_pbr, dfTukey_within_pbc) %>%
  # use group1.treatment as treatment column to facet by in plot
  mutate(treatment = group1.treatment) %>%
  # make comparison column of species 1 vs species 2: estimate comes from group 2-group 1
  mutate(comparison = paste(group2.species, "vs", group1.species, sep = " ")) %>%
  rename(Estimate = estimate) %>%
  mutate(p.adj.signif = ifelse(p.adj.signif == "ns", "", p.adj.signif))

```

#### Plot
```{r heat-map-tukey-treatment-species-within-plot}
# plot
HeatMap_Tukey_Treat_within <- dfTukey_Treat_within %>%
  ggplot(aes(x = dependent_variable_names,
             y = comparison,
             fill = Estimate)) +
  geom_tile(color = "white",
            lwd = 0.5,
            linetype = 1) +
  coord_fixed() +
  facet_wrap(~treatment) +
  geom_text(aes(label = paste(p.adj.signif, round(Estimate, digits = 4), sep = " ")), 
            color = "white",
            size = 3) +
  scale_x_discrete(labels = c("Pigmentation (RGB)", expression(paste("Surface Area (c", m^3, ")", sep = "")), "Weight (g)")) +
  scale_fill_stepsn(
    name = "Estimate",
    colors = cols_rocket
  ) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_blank(),
        axis.text.x = element_text(angle = 45, 
                                   hjust = 1, 
                                   color = "black",
                                   size = 9),
        axis.text.y = element_text(color = "black", face = "italic", size = 9),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(size = 9),
        legend.title = element_text(size = 9))

HeatMap_Tukey_Treat_within

#save plot
# ggsave("Figures/StatisticsResults/HeatMaps/HeatMap-TukeyTest-TreatmentSpecies-Within.png",
#        plot = HeatMap_Tukey_Treat_within,
#        width = 9, height = 4)

```


# Box Plots 
## Tukey Results
### Treatment and Species
Create a boxplot of the differences between treatments for all response variables and faceted by species. This looks like 9 panels: 3 columns by 3 rows. Each column is a different species. Each row is a different response variable. All 3 treatments are shown in each panel. 
```{r boxplot-all-vars-treat}
# long format dfCoral
dfCoral_long <- dfCoral %>%
  pivot_longer(cols = 5:7,
               names_to = "var_names", 
               values_to = "var_vals") %>%
  mutate(var_names = case_when(
    var_names == "weight_g_weekly" ~ "Weight (g)",
    var_names == "surface_area_cm3_weekly" ~ "Surface Area (cm^3)",
    var_names == "rgb_weekly" ~ "Pigmentation (RGB)"
  )) 

# can't find a way to edit the superscript on surface area ylab, so will have to do that in post
BoxPlot_all_treat_species <- 
  ggplot(data = dfCoral_long,
         aes(x = Treatment,
             y = var_vals,
             fill = Treatment)) +
  geom_boxplot(outlier.alpha = 0.2) +
  scale_fill_manual(values = cols_rocket_sm) +
  facet_grid(rows = vars(var_names),
             cols = vars(Species),
             scales = "free",
             switch = "y") +
  geom_signif(comparisons=list(c("Control", "PBC"), 
                               c("Control", "PBR"), 
                               c("PBC", "PBR")),
              map_signif_level = TRUE,
              step_increase = 0.06,
              tip_length = 0) +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line = element_blank(),
        axis.text.x = element_text(color = "black", size = 15),
        axis.text.y = element_text(color = "black", size = 15),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        strip.background = element_rect(fill = "white"),
        strip.text.x = element_text(face = "italic", size = 15),
        strip.text.y = element_text(size = 15),
        strip.background.y = element_rect(size = 0),
        strip.placement = "outside",
        legend.text = element_text(size = 15), 
        legend.title = element_text(size = 15))

BoxPlot_all_treat_species

# save plot 
# ggsave("Figures/StatisticsResults/BoxPlots/BoxPlot-All-Treatment-Species.png",
#       plot = BoxPlot_all_treat_species,
#       height = 13, width = 9)

```


## Daily Light Integral (DLI) PAR Measurements 
Create a boxplot for supplemental materials that shows the daily light integral for all days in each trial. This data is taken from the LiCor dataloggers, and the validity of the data cannot be totally verified unfortunately. Steady PAR measurements mentioned in paper are verified using the PARwise sensor. 
```{r box-plot-dli}
# dli per trial, color by treatment 
boxplot_DLI <- dfEnvVar_long %>%
  filter(variable_names == "DLI") %>%
  filter(Trial != "Control") %>%
  mutate(Trial = factor(Trial, levels = c("PBR 1", "PBC 1", "PBR 2", "PBC 2", "PBR 3", "PBC 3", "PBR 4"))) %>%
  ggplot(aes(x = Trial,
             y = variable_values,
             fill = Treatment)) +
  geom_boxplot() +
  labs(y = expression(paste("Daily Light Integral (mol•",m^-2,"•",d^-1,")", sep = "")))+
  scale_fill_manual(values = cols_rocket_sm[2:3]) +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        panel.border = element_blank(),
        axis.line.x.bottom = element_line(),
        axis.line.y.left = element_line(),
        axis.text.x = element_text(color = "black", size = 12),
        axis.text.y = element_text(color = "black", size = 12),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 12),
        legend.text = element_text(size = 12), 
        legend.title = element_text(size = 12))

boxplot_DLI

# save plot 
# ggsave("Figures/ConfoundingFactors/BoxPlot-DLI-Per-Trial.png",
#        plot = boxplot_DLI,
#        height = 4)

```

## PBC Trials: RGB
Create a boxplot of RGB values for the PBC trials for the supplemental materials because all 3 trials were found to be significantly different from each other in the statistical testing. This figure shows where we conclude the differences come from: differences in RGB. 
```{r box-plot-pbc-trials-rgb}
boxplot_PBCTrials_RGB <- dfCoral_long %>%
  filter(Treatment == "PBC" & var_names == "Pigmentation (RGB)") %>%
  ggplot(aes(x = Trial,
             y = var_vals, 
             fill = Trial)) +
  geom_boxplot() +
  scale_fill_manual(values = cols_rocket) +
  labs(y = "Pigmentation (RGB)") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        panel.border = element_blank(),
        axis.line.x.bottom = element_line(),
        axis.line.y.left = element_line(),
        axis.text.x = element_text(color = "black", size = 12),
        axis.text.y = element_text(color = "black", size = 12),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 12),
        legend.text = element_text(size = 12), 
        legend.title = element_text(size = 12))

boxplot_PBCTrials_RGB

# save plot 
# ggsave("Figures/StatisticsResults/BoxPlots/BoxPlot-RGB-PBCTrials.png",
#        plot = boxplot_PBCTrials_RGB)

```

# Tables
## Confounding Factors 
### Ranges: Stated
Make a table of the range of each trial for all confounding factors. This will look like min - max for each variable. 
```{r table-counfounding-factors-ranges-stated}
# min - max
dfConf_range <- dfEnvVar %>%
  select(-`Nitrate HR`) %>%
  mutate(Trial = factor(Trial, levels = c("PBR 1", "PBC 1", "PBR 2", "PBC 2", "PBR 3", "PBC 3", "Control"))) %>%
  group_by(Trial) %>%
  summarise(`Temperature (ºC)` = paste(round(min(`Temperature (ºC)`, 
                                                 na.rm = T), digits = 2), 
                                       round(max(`Temperature (ºC)`, 
                                                 na.rm = T), digits = 2),
                                       sep = " - "), 
            `pH` = paste(round(min(pH, na.rm = T), digits = 2), 
                       round(max(pH, na.rm = T), digits = 2),
                       sep = " - "),
            `Salinity (PSU)` = paste(round(min(`Salinity (PSU)`, 
                                               na.rm = T), digits = 2), 
                                     round(max(`Salinity (PSU)`, 
                                               na.rm = T), digits = 2),
                                     sep = " - "),
            `ORP (mV)` = paste(round(min(`ORP (mV)`, na.rm = T), digits = 2), 
                               round(max(`ORP (mV)`, na.rm = T), digits = 2),
                               sep = " - "),
            `Chlorophyll (RFU)` = paste(round(min(`Chlorophyll (RFU)`, 
                                                  na.rm = T), digits = 2), 
                                        round(max(`Chlorophyll (RFU)`, 
                                                  na.rm = T), digits = 2),
                                        sep = " - "),
            `Phycoerythrin (RFU)` = paste(round(min(`Phycoerythrin (RFU)`, 
                                                    na.rm = T), digits = 2), 
                                          round(max(`Phycoerythrin (RFU)`, 
                                                    na.rm = T), digits = 2),
                                          sep = " - "),
            `Alkalinity` = paste(round(min(`Alkalinity`, na.rm = T), digits = 2),
                                 round(max(`Alkalinity`, na.rm = T), digits = 2),
                                 sep = " - "),
            `DLI` = paste(round(min(`DLI`, na.rm = T), digits = 2), 
                          round(max(`DLI`, na.rm = T), digits = 2),
                          sep = " - "),
            `Ammonia` = paste(round(min(`Ammonia`, na.rm = T), digits = 2), 
                              round(max(`Ammonia`, na.rm = T), digits = 2),
                              sep = " - "),
            `Calcium` = paste(round(min(`Calcium`, na.rm = T), digits = 2), 
                              round(max(`Calcium`, na.rm = T), digits = 2),
                              sep = " - "),
            `Nitrate MR` = paste(round(min(`Nitrate MR`, na.rm = T), digits = 2),
                                 round(max(`Nitrate MR`, na.rm = T), digits = 2),
                                 sep = " - "),
            `Phosphate` = paste(round(min(`Phosphate`, na.rm = T), digits = 2), 
                                round(max(`Phosphate`, na.rm = T), digits = 2),
                                sep = " - "),
            `Silica` = paste(round(min(`Silica`, na.rm = T), digits = 2), 
                             round(max(`Silica`, na.rm = T), digits = 2),
                             sep = " - "),
            `Iron` = paste(round(min(`Iron`, na.rm = T), digits = 2), 
                           round(max(`Iron`, na.rm = T), digits = 2),
                           sep = " - "),
            `Turbidity` = paste(round(min(`Turbidity`, na.rm = T), digits = 2), 
                                round(max(`Turbidity`, na.rm = T), digits = 2),
                                sep = " - ")
  ) %>%
  mutate(Alkalinity = ifelse(Alkalinity == "Inf - -Inf", NA, Alkalinity),
         DLI = ifelse(DLI == "Inf - -Inf", NA, DLI),
         Silica = ifelse(Silica == "Inf - -Inf", NA, Silica),
         Iron = ifelse(Iron == "Inf - -Inf", NA, Iron)) %>%
  rename(`Alkalinity (mg/L)` = `Alkalinity`,
         `Ammonia (mg/L)` = `Ammonia`, 
         `Calcium (mg/L)` = `Calcium`, 
         `Nitrate (mg/L)` = `Nitrate MR`, 
         `Phosphate (mg/L)` = `Phosphate`,
         `Silica (mg/L)` = `Silica`, 
         `Iron (mg/L)` = `Iron`, 
         `Turbidity (FAU)` = `Turbidity`)

write_xlsx(dfConf_range, "Figures/ConfoundingFactors/ConfoundingFactors-Ranges-Stated.xlsx")
 
```

### Ranges: Difference
Make a table of the range of each trial for all confounding factors. This will look like 'difference' (of min and max) for each variable. 
```{r table-counfounding-factors-ranges-difference}
# difference
dfConf_range_diff <- dfEnvVar %>%
  select(-`Nitrate HR`) %>%
  mutate(Trial = factor(Trial, levels = c("PBR 1", "PBC 1", "PBR 2", "PBC 2", "PBR 3", "PBC 3", "Control"))) %>%
  group_by(Trial) %>%
  summarise(`Temperature (ºC)` = round(max(`Temperature (ºC)`, na.rm = T) - min(`Temperature (ºC)`, na.rm = T), digits = 2),
            `pH` = round(max(`pH`, na.rm = T) - min(`pH`, na.rm = T), digits = 2),
            `Salinity (PSU)` = round(max(`Salinity (PSU)`, na.rm = T) - min(`Salinity (PSU)`, na.rm = T), digits = 2),
            `ORP (mV)` = round(max(`ORP (mV)`, na.rm = T) - min(`ORP (mV)`, na.rm = T), digits = 2),
            `Chlorophyll (RFU)` = round(max(`Chlorophyll (RFU)`, na.rm = T) - min(`Chlorophyll (RFU)`, na.rm = T), digits = 2),
            `Phycoerythrin (RFU)` = round(max(`Phycoerythrin (RFU)`, na.rm = T) - min(`Phycoerythrin (RFU)`, na.rm = T), digits = 2),
            `Alkalinity` = round(max(`Alkalinity`, na.rm = T) - min(`Alkalinity`, na.rm = T), digits = 2),
            `DLI` = round(max(`DLI`, na.rm = T) - min(`DLI`, na.rm = T), digits = 2),
            `Ammonia` = round(max(`Ammonia`, na.rm = T) - min(`Ammonia`, na.rm = T), digits = 2),
            `Calcium` = round(max(`Calcium`, na.rm = T) - min(`Calcium`, na.rm = T), digits = 2),
            `Nitrate MR` = round(max(`Nitrate MR`, na.rm = T) - min(`Nitrate MR`, na.rm = T), digits = 2),
            `Phosphate` = round(max(`Phosphate`, na.rm = T) - min(`Phosphate`, na.rm = T), digits = 2),
            `Silica` = round(max(`Silica`, na.rm = T) - min(`Silica`, na.rm = T), digits = 2),
            `Iron` = round(max(`Iron`, na.rm = T) - min(`Iron`, na.rm = T), digits = 2),
            `Turbidity` = round(max(`Turbidity`, na.rm = T) - min(`Turbidity`, na.rm = T), digits = 2)
  ) %>%
  mutate(Alkalinity = ifelse(Alkalinity == "-Inf", NA, Alkalinity),
         DLI = ifelse(DLI == "-Inf", NA, DLI),
         Silica = ifelse(Silica == "-Inf", NA, Silica),
         Iron = ifelse(Iron == "-Inf", NA, Iron)) %>%
  rename(`Alkalinity (mg/L)` = `Alkalinity`,
         `Ammonia (mg/L)` = `Ammonia`, 
         `Calcium (mg/L)` = `Calcium`, 
         `Nitrate (mg/L)` = `Nitrate MR`, 
         `Phosphate (mg/L)` = `Phosphate`,
         `Silica (mg/L)` = `Silica`, 
         `Iron (mg/L)` = `Iron`, 
         `Turbidity (FAU)` = `Turbidity`)

write_xlsx(dfConf_range_diff, "Figures/ConfoundingFactors/ConfoundingFactors-Ranges-Difference.xlsx")

```



# References
Hackerott et al. 2025: <https://doi.org/10.21203/rs.3.rs-5247850/v1>
ColorBrewer: <https://colorbrewer2.org/#>
