---
Title: "Mini Ocean Lighting Project Wrangling: Compounding Factors"
Collaborators: "Renee Grambihler, Savanna Saunders"
Project: "Mini Ocean Light Data"
Institution: "Biosphere 2 Ocean"
Description: "Wrangles environmental data for use in compounding factors statistical testing"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---
**Input**
  `Mini Ocean Nutrient Data.xlsx`: colorimetry data for mini ocean for summer, fall, and spring
  
  `Raceway2NutrientData.xlsx`: colorimetry data for all sites for summer, fall, and spring
  
  `PRO DSS MO All - Sheet1.csv`: water quality data from YSI ProDSS instrument for mini ocean for summer, fall, and spring
  
  `prodssbackup-250328-140416 All Sites.csv`: water quality data from YSI ProDSS instrument for all sites from September 2024 - March 2025. Used to get Raceway 2 (Control) parameters. 
  
  `Raceway2Parameters_2024-05-29_2025-09-30_RoundsForm.xlsx`: water quality data from Google Form of daily rounds for Raceway 2 (Control) to complete time gap with summer data. YSI ProDSS is used to obtain data, but fewer parameters are recorded: only temperature (F), salinity (PSU), and pH. 
  
  `DLI-All.xlsx`: daily light integral data made from PAR data recorded over summer, fall, and spring. File created in `Wrangling-PAR.Rmd`. 

**Output**
  `All-Environmental-Variables-Wide.xlsx`: all available data from all environmental variables recorded over summer, fall, and spring experiments in wide form.
  
  `All-Environmental-Variables-Long.xlsx`: all available data from all environmental variables recorded over summer, fall, and spring experiments in long form. 

## Import Libraries
```{r import-libraries}
library(tidyverse)
library(readxl)
library(writexl)
library(rstatix)
library(lubridate)
# get rid of scientific notation
options(scipen = 9999)
# avoid dplyr error message
options(dplyr.summarise.inform = FALSE)  
```

# Import Data
```{r import-data}
#### Nutrients #### 
# mini ocean (colorimetry) 
dfNutrients <- read_excel("InputData/Raw/WaterQuality/Nutrients/Mini Ocean Nutrient Data.xlsx") %>%
  select(-Initials, -Notes, -Location)
   
# raceway 2 (colorimetry)
dfNutrients_r2 <- read_excel("InputData/Raw/WaterQuality/Nutrients/Raceway2NutrientData.xlsx") %>%
  #grab just raceway 2 obs
  filter(Location == "Raceway 2") %>%
  #get rid of location column to match other nutrients df
  select(-Location)
 
#### Water Quality #### 
# YSI ProDSS data from Mini Ocean for 6/3/2024 - 3/17/2025
dfYSI <- dfYSI <- read_csv("InputData/Raw/WaterQuality/ProDSS/PRO DSS MO All - Sheet1.csv") %>%
  select(-`User ID`, -`Unit ID`)%>%
  mutate(Date = as.Date(Date, format = "%m/%d/%y"))%>%
  rename(`Temperature (ºF)` = "¬∞F",
         `Dissolved Oxygen (%)` = "DO %",
         `Dissolved Oxygen (mg/L)` = "DO mg/L",
         `Conductivity (µS/cm)` = "C-uS/cm",
         `Salinity (PSU)` = "SAL-PSU",
         `ORP (mV)` = "ORP mV",
         `Chlorophyll (RFU)` = "Chl RFU",
         `Phycoerythrin (RFU)` = "PE RFU")

# Control Values: 
# YSI ProDSS data from all site for Sept 2024 - March 2025 (to get Control from R2)
dfYSI_allsites <- read_csv("InputData/Raw/WaterQuality/ProDSS/prodssbackup-250328-140416 All Sites.csv") %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%y"))%>%
  # take out unit id, user id, dissolved oxygen, and conductivity columns
  select(-c("Unit ID", "User ID", "DO %", "DO mg/L", "C-uS/cm")) %>%
  # grab just Raceway 2 data
  filter(Site == "R2") %>%
  rename(`Temperature (ºF)` = "F",
         `Salinity (PSU)` = "SAL-PSU",
         `ORP (mV)` = "ORP mV",
         `Chlorophyll (RFU)` = "Chl RFU",
         `Phycoerythrin (RFU)` = "PE RFU")

# YSI ProDSS R2 Basic Parameters May 2024 - Sept 2024 (to get control. only have temp, salinity, and pH for Connor's trials because ProDSS downloading port was broken by the time we needed to add Connor's data to Savanna's honors thesis data)
dfYSI_basic <- read_excel("InputData/Raw/WaterQuality/ProDSS/Raceway2Parameters_2024-05-29_2025-09-30_RoundsForm.xlsx") %>%
  separate_wider_delim(cols = Datetime, 
                       delim = " ", 
                       names = c("Date", "Time"))

####  DLI #### 
dfDLI <- read_excel("InputData/Clean/EnvironmentalData/DLI-All.xlsx") %>%
  #remove sum of par values for each day which was used to make DLI values
  select(-daily_par_sum)

```

# Data Wrangling
Using the mean of each day where there is more than one observation for each day. 
## Nutrient Data
```{r data-wrangling-nutrient}
# mini ocean data 
dfNutrients_edit <- dfNutrients %>%
  mutate(Ammonia = ifelse(Ammonia == "---", 0.00, Ammonia), #change --- values to 0.00 for ammonia
         Ammonia = as.numeric(Ammonia),
         #round columns to correct sig figs from colorimeter
         Ammonia = round(Ammonia, digits = 2),
         Calcium = ifelse(Calcium == 2.2, 220, Calcium), #change probable Calcium input error from 2.20 to 220, which matches the range of surrounding data
         Iron = round(Iron, digits = 3),
         Silica = ifelse(Silica == "---", 0.00, Silica), #change --- values to 0.00 for Silica
         Silica = as.numeric(Silica),
         #round columns to correct sig figs from colorimeter
         Silica = round(Silica, digits = 2),
         `Nitrate HR` = round(`Nitrate HR`, digits = 1),
         `Nitrate MR` = round(`Nitrate MR`, digits = 1),
         Phosphate = round(Phosphate, digits = 2)) %>%
  select(-pH) %>% #use pH from YSI since it has more data and is more accurate
  mutate(Date = as.Date(Date)) %>%
   #add trial by dates
  mutate(Trial = NA,
         Trial = ifelse(Date >= "2024-06-03" & Date < "2024-06-17", "PBR 1", Trial),
         Trial = ifelse(Date >= "2024-06-17" & Date < "2024-07-09", "PBC 1", Trial),
         Trial = ifelse(Date >= "2024-07-09" & Date < "2024-07-29", "PBR 2", Trial),
         Trial = ifelse(Date >= "2024-09-29" & Date < "2024-11-11", "PBR 3", Trial),
         Trial = ifelse(Date >= "2024-11-11" & Date < "2024-12-20", "PBC 2", Trial),
         Trial = ifelse(Date >= "2024-12-20" & Date < "2025-02-03", "PBR 4", Trial),
         Trial = ifelse(Date >= "2025-02-03" & Date <= "2025-03-17", "PBC 3", Trial), 
         .after = Date) %>%
  filter(is.na(Trial) == FALSE) %>%
  mutate(Trial = factor(Trial, levels = c("PBR 1", "PBC 1", "PBR 2", "PBR 3", "PBC 2", "PBR 4", "PBC 3"))) %>%
  #use trial column to make treatment column
  separate(col = Trial,
           sep = " ",
           into = c("Treatment", "trial_num"), 
           remove = FALSE) %>%
  select(-trial_num) %>%
  filter(Trial != "PBR 4") #took pbr 4 out of nonparametric due to outliers in dependent variables

# r2 data 
dfNutrients_r2_edit <- dfNutrients_r2 %>%
  mutate(Ammonia = ifelse(Ammonia == "---", 0.00, Ammonia), #change --- values to 0.00 for ammonia
         Ammonia = as.numeric(Ammonia),
         #round columns to correct sig figs from colorimeter
         Ammonia = round(Ammonia, digits = 2),
         Calcium = ifelse(Calcium == 2.2, 220, Calcium), #change probable Calcium input error from 2.20 to 220, which matches the range of surrounding data
         Iron = round(Iron, digits = 3),
         Silica = ifelse(Silica == "---", 0.00, Silica), #change --- values to 0.00 for Silica
         Silica = as.numeric(Silica),
         #round columns to correct sig figs from colorimeter
         Silica = round(Silica, digits = 2),
         `Nitrate HR` = round(`Nitrate HR`, digits = 1),
         `Nitrate MR` = round(`Nitrate MR`, digits = 1),
         Phosphate = round(Phosphate, digits = 2)) %>%
  select(-pH) %>% #use pH from YSI since it has more data and is more accurate
  mutate(Date = as.Date(Date)) %>%
  mutate(Trial = "Control", .after = Date) %>%
  mutate(Treatment = "Control", .after = Trial)
  
dfNutrients_edit <- rbind(dfNutrients_edit, dfNutrients_r2_edit)

```

## YSI ProDSS Data
```{r data-wrangling-prodss}
dfYSI_edit <- dfYSI %>%
  #get rid of columns that won't be used. DO not reliable on prodss
  select(-c("Site", "Dissolved Oxygen (%)", "Dissolved Oxygen (mg/L)", "Conductivity (µS/cm)")) %>%
  #convert fahrenheit to celcius
  mutate(`Temperature (ºC)` = (`Temperature (ºF)` - 32) * 5 / 9, .after = Time) %>%
  select(-`Temperature (ºF)`) %>%
  na.omit() %>%
  group_by(Date) %>%
  #data was recorded once per day except for during scheduled holidays when it was automatically recorded once every 15 minutes, so take average of each day which will reduce multiple days to 1 entry each
  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE))) %>%
  #add trial by dates
  mutate(Trial = case_when(
    Date >= "2024-06-03" & Date < "2024-06-17" ~ "PBR 1",
    Date >= "2024-06-17" & Date < "2024-07-09" ~ "PBC 1",
    Date >= "2024-07-09" & Date < "2024-07-29" ~ "PBR 2",
    Date >= "2024-09-30" & Date < "2024-11-11" ~ "PBR 3",
    Date >= "2024-11-11" & Date < "2024-12-20" ~ "PBC 2",
    Date >= "2024-12-20" & Date < "2025-02-03" ~ "PBR 4",
    Date >= "2025-02-03" & Date <= "2025-03-17" ~ "PBC 3",
  ), .after = Date) %>%
  #use trial column to make treatment column
  separate(col = Trial,
           sep = " ",
           into = c("Treatment", "trial_num"), 
           remove = FALSE) %>%
  select(-trial_num) %>%
  filter(is.na(Trial) == FALSE) %>%
  mutate(Trial = factor(Trial, levels = c("PBR 1", "PBC 1", "PBR 2", "PBR 3", "PBC 2", "PBR 4", "PBC 3"))) %>%
  #filter(Trial != "PBR 4") %>% #took pbr 4 out of nonparametric due to outliers
  filter(Date <= "2025-03-17" & Date >= "2024-06-03") %>%
  filter(Trial != "PBR 4") #took pbr 4 out of nonparametric due to outliers in dependent variables

# ProDSS R2 - Savanna's data 
dfYSI_sav <- dfYSI_allsites %>%
  mutate(Date = as.Date(Date)) %>%
  # filter out dates after 3/17/25
  filter(Date <= "2025-03-17") %>%
  #convert fahrenheit to celcius
  mutate(`Temperature (ºC)` = (`Temperature (ºF)` - 32) * 5 / 9, .after = Time) %>%
  select(-`Temperature (ºF)`) %>%
  na.omit() %>%
  group_by(Date) %>%
  #data was recorded once per day except for a few days so take average of each day which will reduce multiple days to 1 entry each
  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE))) %>%
  mutate(Trial = "Control", .after = Date) %>%
  mutate(Treatment = "Control", .after = Trial)

# ProDSS R2 - Connor's data
dfYSI_basic_edit <- dfYSI_basic %>%
  mutate(Date = as.Date(Date)) %>%
  #convert fahrenheit to celcius
  mutate(`Temperature (ºC)` = (temp_f - 32) * 5 / 9, .after = Time) %>%
  select(-temp_f) %>%
  na.omit() %>%
  group_by(Date) %>%
  #data was recorded once per day except for a few days so take average of each day which will reduce multiple days to 1 entry each
  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE))) %>%
  mutate(Trial = "Control", .after = Date) %>%
  mutate(Treatment = "Control", .after = Trial) %>%
  rename(`Salinity (PSU)` = "salinity") %>%
  filter(Date < "2024-09-20") %>%
  # move column to bind with rest of data later
  relocate(pH, .after = `Salinity (PSU)`) %>%
  # add empty columns to bind with rest of data later
  mutate(`ORP (mV)` = NA, 
         `Chlorophyll (RFU)` = NA,
         `Phycoerythrin (RFU)` = NA)

dfYSI_r2 <- rbind(dfYSI_basic_edit, dfYSI_sav)

```

## All Environmental Data
```{r combine-data}
dfAllEnvVar <- data.frame(Date = seq.Date(from = as.Date("2024-05-29"), to = as.Date("2025-03-17"), by = "1 day")) %>%
  mutate(Trial = case_when(
    Date >= "2024-06-03" & Date < "2024-06-17" ~ "PBR 1",
    Date >= "2024-06-17" & Date < "2024-07-09" ~ "PBC 1",
    Date >= "2024-07-09" & Date < "2024-07-29" ~ "PBR 2",
    Date >= "2024-09-30" & Date < "2024-11-11" ~ "PBR 3",
    Date >= "2024-11-11" & Date < "2024-12-20" ~ "PBC 2",
    Date >= "2025-02-03" & Date <= "2025-03-17" ~ "PBC 3")) %>%
  left_join(dfYSI_edit) %>%
  rbind(dfYSI_r2) %>%
  left_join(dfDLI) %>%
  left_join(dfNutrients_edit)%>%
  filter(is.na(Trial) == FALSE) %>%
  select(-Treatment) %>%
  separate(col = Trial,
           sep = " ",
           into = c("Treatment", "trial_num"),
           remove = FALSE) %>%
  select(-trial_num) %>%
  filter(is.na(Trial) == FALSE) %>%
  # add DLI PBR2 value - only one day was recorded and it got lost in the wrangling
  mutate(DLI = ifelse(Trial == "PBR 2", 0.10036601000, DLI)) 
                   

# long version of dfAllEnvVar
dfAllEnvVar_long <- dfAllEnvVar %>%
  pivot_longer(cols = 4:19,
               names_to = "variable_names", 
               values_to = "variable_values")

```


```{r write-excel-files}
# environmental variables wide
write_xlsx(dfAllEnvVar, "InputData/Clean/EnvironmentalData/All-Environmental-Variables-Wide.xlsx")

# environmental variables long
write_xlsx(dfAllEnvVar_long, "InputData/Clean/EnvironmentalData/All-Environmental-Variables-Long.xlsx")

```

