---
Title: "Mini Ocean Lighting Project Wrangling: PAR Data and DLI Pipeline"
Collaborators: "Connor Fenn, Savanna Saunders, Renee Grambihler"
Project: "Mini Ocean Light Data"
Institution: "Biosphere 2 Ocean"
Description: "Wrangles PAR data obtained from LICOR instruments in the Mini Ocean during the summer, fall, and spring 2024-2025 Coral Lighting Experiments. Combines all PAR measurements from across all trials and then calculates the Daily Light Integral (DLI) for each day to use in downstream statistics."
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

**Input**
  `MO-PBR1.TXT`: output file from LiCor LI-1500 Light Sensor Logger connected to LI-192 Underwater Quantum Sensor to record PAR for Treatment PBR Trial 1: 05/29/2024 - 06/17/2024
        
  `MO-PBR2.TXT`: output file from LiCor LI-1500 Light Sensor Logger connected to LI-192 Underwater Quantum Sensor to record PAR for Treatment PBR Trial 2: 07/09/2024 - 07/29/2024
        
  `MO-PBR3.TXT`: output file from LiCor LI-1500 Light Sensor Logger connected to LI-192 Underwater Quantum Sensor to record PAR for Treatment PBR Trial 3: 09/30/2024 - 11/11/2024
        
  `MO-PBR4.TXT`: output file from LiCor LI-1500 Light Sensor Logger connected to LI-192 Underwater Quantum Sensor to record PAR for Treatment PBR Trial 4: 12/20/2024 - 02/03/2025
        
  `MO-PBC1.TXT`: output file from LiCor LI-1500 Light Sensor Logger connected to LI-192 Underwater Quantum Sensor to record PAR for Treatment PBC Trial 1: 06/17/2024 - 07/09/2024
        
  `MO-PBC2.TXT`: output file from LiCor LI-1500 Light Sensor Logger connected to LI-192 Underwater Quantum Sensor to record PAR for Treatment PBC Trial 2: 11/11/2024 - 12/20/2024
        
  `MO-PBC3.TXT`: output file from LiCor LI-1500 Light Sensor Logger connected to LI-192 Underwater Quantum Sensor to record PAR for Treatment PBC Trial 3: 02/03/2025 - 03/17/2025
  
**Output**
  `AllPAR.xlsx`: spreadsheet of all the PAR values from each trial.
        
  `DLI-All.xlsx`: spreadsheet of daily light integral values for each day from each trial.
  

# Import Libraries
```{r Libraries, include=FALSE}
library(tidyverse)
library(lubridate)
library(data.table)
library(tibbletime)
library(ggspectra)
library(hms)
library(stringr)
library(writexl)
#opts_chunk$set(echo = TRUE)
```

# Import Data
Text files will need to be edited slightly before they are read in so that the heading they come with is deleted except for the column names. Files here should be already edited. 
```{r Import and format data}
# PBR-1
dfPAR_PBR1 <- read_tsv("InputData/Raw/PAR/MO-PBR1.TXT",
                   col_names = TRUE)

# PBR-2
dfPAR_PBR2 <- read_tsv("InputData/Raw/PAR/MO-PBR2.TXT",
                   col_names = TRUE)
                   
# PBR-3
dfPAR_PBR3 <- read_tsv("InputData/Raw/PAR/MO-PBR3.TXT",
                   col_names = TRUE)
                   
# PBR-4
dfPAR_PBR4 <- read_tsv("InputData/Raw/PAR/MO-PBR4.TXT",
                   col_names = TRUE)

# PBC-1
dfPAR_PBC1 <- read_tsv("InputData/Raw/PAR/MO-PBC1.TXT",
                   col_names = TRUE)

# PBC-2
dfPAR_PBC2 <- read_tsv("InputData/Raw/PAR/MO-PBC2.TXT",
                   col_names = TRUE)

# PBC-3
dfPAR_PBC3 <- read_tsv("InputData/Raw/PAR/MO-PBC3.TXT",
                   col_names = TRUE)

```

# Data Wrangling
The LI-CORE files got messed up at some point and many entries moved the columns to the left by 1, which puts a bunch of NAs in when trying to do downstream calculations and ultimately removes a ton of data accidentally. Detected this by using `head(which(is.na(dfPAR_PBR1), arr.ind=TRUE))` after wrangling was done. Also a reason to not do this step until after importing the files in their original state with as few changes as possible

## PBR 1
```{r data-wrangling-pbr1}
# select only columns we want: Date, Time, PAR values
# rename column with PAR values 
# make sure there aren't any repeat rows
# change column classes to ones we want: Date is date class, PAR is numeric class, Time will stay character until it gets messed with in later calculations due to there being more than 1 observation per minute in most cases 
# create Trial column with the lighting trial from that data file

# remove column 7 because it's an extra column from importing
dfPAR_PBR1_edit <- dfPAR_PBR1 %>%
  select(-`...7`) 

#check for NA values once empty column is gone
head(which(is.na(dfPAR_PBR1_edit), arr.ind=TRUE)) 
# row 47803, columns 4 and 5
# rows 21792 - 21795 column 6

# double check for columns being moved over using str_detect(Record, "-") to see if there are dashes used in the Date column in the Record column
dfPAR_PBR1_edit %>%
 filter(str_detect(Record, "-") == TRUE) %>%
  nrow()
# 66490 rows 

# Keep regularly formatted data in this dataframe
dfPAR_PBR1_edit <- dfPAR_PBR1_edit %>%
  filter(str_detect(DATAH, "DATA") == TRUE) %>%
  select(-Record, -CHK, -DATAH) %>%
  rename(PAR = MINIOCEA)
#check for NA values
head(which(is.na(dfPAR_PBR1_edit), arr.ind=TRUE)) #none

# keep irregularly formatted data in new data frame
dfPAR_PBR1_filter <- dfPAR_PBR1 %>%
  select(-`...7`) %>%
  filter(str_detect(Record, "-") == TRUE) %>%
  select(-MINIOCEA, -CHK, -DATAH) %>%
  rename(PAR = Time) %>%
  rename(Time = Date) %>%
  rename(Date = Record)
#check for NA values
head(which(is.na(dfPAR_PBR1_filter), arr.ind=TRUE)) #one entry which looks like a mistake by the logger

dfPAR_PBR1_bind <- rbind(dfPAR_PBR1_edit, dfPAR_PBR1_filter) %>%
  mutate(Trial = "PBR 1", .after = Time)

#check for NA values
head(which(is.na(dfPAR_PBR1_bind), arr.ind=TRUE)) #same entry as before

# get rid of row with na 
dfPAR_PBR1_final <- dfPAR_PBR1_bind %>%
  na.omit()%>%
  mutate(Date = as.Date(Date)) %>%
  mutate(Time = gsub("\\..*","",Time)) %>%
  # make sure PAR values are numeric
  mutate(PAR = as.numeric(PAR)) %>%
  # mean PAR value per minute
  mutate(Time = gsub('.{3}$', ':00', Time)) %>%
  filter(PAR != 0)  %>% # PAR value of 0 with no decimal places seems to be some kind of logging error since all the other values have decimal places, and many of them are during daylight hours
  group_by(Date, Time, Trial) %>%
  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE))) 


```

## PBR 2
```{r data-wrangling-pbr2}
# select only columns we want: Date, Time, PAR values
# rename column with PAR values 
# make sure there aren't any repeat rows
# change column classes to ones we want: Date is date class, PAR is numeric class, Time will stay character until it gets messed with in later calculations due to there being more than 1 observation per minute in most cases 
# create Trial column with the lighting trial from that data file

# remove column 7
dfPAR_PBR2_edit <- dfPAR_PBR2 %>%
  select(-`...7`)

#check for NA values
head(which(is.na(dfPAR_PBR2_edit), arr.ind=TRUE)) #none
# double check for columns being moved over using str_detect(Record, "-") to see if there are dashes used in the Date column in the Record column
dfPAR_PBR2_edit %>%
 filter(str_detect(Record, "-") == TRUE) %>%
  nrow()
# 0 rows 
# irregularities from PBR1 not here

# finish normal data wrangling
dfPAR_PBR2_final <- dfPAR_PBR2_edit %>%
  select(-Record, -CHK, -DATAH) %>%
  rename(PAR = MINIOCEA) %>%
  mutate(Trial = "PBR 2", .after = Time)%>%
  mutate(Date = as.Date(Date)) %>%
  # make sure PAR values are numeric
  mutate(PAR = as.numeric(PAR)) %>%
  # mean PAR value per minute
  mutate(Time = gsub('.{3}$', ':00', Time)) %>%
  filter(PAR != 0)  %>% # PAR value of 0 with no decimal places seems to be some kind of logging error since all the other values have decimal places, and many of them are during daylight hours
  group_by(Date, Time, Trial) %>%
  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE))) 

#check if time has milliseconds 
dfPAR_PBR2_final %>%
  filter(str_detect(Time, "\\..*") == TRUE) %>%
  nrow() #none

#check for NA values
head(which(is.na(dfPAR_PBR2_final), arr.ind=TRUE)) #still none

```

## PBR 3
```{r data-wrangling-pbr3}
# select only columns we want: Date, Time, PAR values
# rename column with PAR values 
# make sure there aren't any repeat rows
# change column classes to ones we want: Date is date class, PAR is numeric class, Time will stay character until it gets messed with in later calculations due to there being more than 1 observation per minute in most cases 
# create Trial column with the lighting trial from that data file

# remove column 7
dfPAR_PBR3_edit <- dfPAR_PBR3 %>%
  select(-`...7`)

#check for NA values
head(which(is.na(dfPAR_PBR3_edit), arr.ind=TRUE)) #none
# double check for columns being moved over using str_detect(Record, "-") to see if there are dashes used in the Date column in the Record column
dfPAR_PBR3_edit %>%
 filter(str_detect(Record, "-") == TRUE) %>%
  nrow()
# 0 rows 
# irregularities from PBR1 not here

# finish normal data wrangling
dfPAR_PBR3_final <- dfPAR_PBR3_edit %>%
  select(-Record, -CHK, -DATAH) %>%
  filter(Date <= "2024-11-11") %>%
  rename(PAR = MO) %>%
  mutate(Trial = "PBR 3", .after = Time)%>%
  mutate(Date = as.Date(Date))

#check if time has milliseconds 
dfPAR_PBR3_final %>%
  filter(str_detect(Time, "\\..*") == TRUE) %>%
  nrow() #none

#check for NA values
head(which(is.na(dfPAR_PBR3_final), arr.ind=TRUE)) #still none

```

## PBR 4
```{r data-wrangling-pbr4}
# select only columns we want: Date, Time, PAR values
# rename column with PAR values 
# make sure there aren't any repeat rows
# change column classes to ones we want: Date is date class, Time is time class, PAR is numeric class
# create datetime column by pasting Date and Time data next to each other in each cell 
# create Trial column with the lighting trial from that data file

# remove column 7
dfPAR_PBR4_edit <- dfPAR_PBR4 %>%
  select(-`...7`)

#check for NA values
head(which(is.na(dfPAR_PBR4_edit), arr.ind=TRUE)) #2 NA's in col 6
# double check for columns being moved over using str_detect(Record, "-") to see if there are dashes used in the Date column in the Record column
dfPAR_PBR4_edit %>%
 filter(str_detect(Record, "-") == TRUE) %>%
  nrow()
# 0 rows 
# irregularities from PBR1 not except for some NAs in column 6 which will be removed anyways

# finish normal data wrangling
dfPAR_PBR4_final <- dfPAR_PBR4_edit %>%
  select(-Record, -CHK, -DATAH) %>%
  rename(PAR = MO) %>%
  mutate(Trial = "PBR 4", .after = Time)%>%
  mutate(Date = as.Date(Date))

#check if time has milliseconds 
dfPAR_PBR4_final %>%
  filter(str_detect(Time, "\\..*") == TRUE) %>%
  nrow() #none

#check for NA values
head(which(is.na(dfPAR_PBR4_final), arr.ind=TRUE)) #none


```
## PBC 1
```{r data-wrangling-pbc1}
# select only columns we want: Date, Time, PAR values
# rename column with PAR values 
# make sure there aren't any repeat rows
# change column classes to ones we want: Date is date class, Time is time class, PAR is numeric class
# create datetime column by pasting Date and Time data next to each other in each cell 
# create Trial column with the lighting trial from that data file

# remove column 7
dfPAR_PBC1_edit <- dfPAR_PBC1 %>%
  select(-`...7`)

#check for NA values
which(is.na(dfPAR_PBC1_edit), arr.ind=TRUE) 
# 7 NAs: 4 in row 61, 2 in row 11662, 1 in row 164. remove instances

# double check for columns being moved over using str_detect(Record, "-") to see if there are dashes used in the Date column in the Record column
dfPAR_PBC1_edit %>%
 filter(str_detect(Record, "-") == TRUE) %>%
  nrow()
# 0 rows 

# finish normal data wrangling
dfPAR_PBC1_final <- dfPAR_PBC1_edit %>%
  na.omit() %>% # remove NA values
  select(-Record, -CHK, -DATAH) %>%
  rename(PAR = MINIOCEA) %>%
  mutate(Trial = "PBC 1", .after = Time)%>%
  mutate(Date = as.Date(Date)) %>%
  # make sure PAR values are numeric
  mutate(PAR = as.numeric(PAR)) %>%
  # mean PAR value per minute
  mutate(Time = gsub('.{3}$', ':00', Time)) %>%
  filter(PAR != 0)  %>% # PAR value of 0 with no decimal places seems to be some kind of logging error since all the other values have decimal places, and many of them are during daylight hours
  group_by(Date, Time, Trial) %>%
  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE))) 

#check if time has milliseconds 
dfPAR_PBC1_final %>%
  filter(str_detect(Time, "\\..*") == TRUE) %>%
  nrow() #none

#check for NA values
head(which(is.na(dfPAR_PBC1_final), arr.ind=TRUE)) #none


```

## PBC 2
```{r data-wrangling-pbc2}
# select only columns we want: Date, Time, PAR values
# rename column with PAR values 
# make sure there aren't any repeat rows
# change column classes to ones we want: Date is date class, Time is time class, PAR is numeric class
# create datetime column by pasting Date and Time data next to each other in each cell 
# create Trial column with the lighting trial from that data file

# remove column 7
dfPAR_PBC2_edit <- dfPAR_PBC2 %>%
  select(-`...7`)

#check for NA values
which(is.na(dfPAR_PBC2_edit), arr.ind=TRUE) #none

# double check for columns being moved over using str_detect(Record, "-") to see if there are dashes used in the Date column in the Record column
dfPAR_PBC2_edit %>%
 filter(str_detect(Record, "-") == TRUE) %>%
  nrow()
# 0 rows 

# finish normal data wrangling
dfPAR_PBC2_final <- dfPAR_PBC2_edit %>%
  select(-Record, -CHK, -DATAH) %>%
  rename(PAR = MO) %>%
  mutate(Trial = "PBC 2", .after = Time)%>%
  mutate(Date = as.Date(Date))

#check if time has milliseconds 
dfPAR_PBC2_final %>%
  filter(str_detect(Time, "\\..*") == TRUE) %>%
  nrow() #none

#check for NA values
head(which(is.na(dfPAR_PBC2_final), arr.ind=TRUE)) #none


```

## PBC 3
Some PBR3 observations got appended to the PBR3 file on the datalogger, so will be fixing that here. Figured this out via plotting downstream.
```{r data-wrangling-pbc3}
# select only columns we want: Date, Time, PAR values
# rename column with PAR values 
# make sure there aren't any repeat rows
# change column classes to ones we want: Date is date class, Time is time class, PAR is numeric class
# create datetime column by pasting Date and Time data next to each other in each cell 
# create Trial column with the lighting trial from that data file

# remove column 7
dfPAR_PBC3_edit <- dfPAR_PBC3 %>%
  select(-`...7`)

#check for NA values
which(is.na(dfPAR_PBC3_edit), arr.ind=TRUE) #two in column 6 which is getting deleted anyways

# double check for columns being moved over using str_detect(Record, "-") to see if there are dashes used in the Date column in the Record column
dfPAR_PBC3_edit %>%
 filter(str_detect(Record, "-") == TRUE) %>%
  nrow()
# 0 rows 

### some PBR3 observations got appended to the PBR3 file on the datalogger: 
dfPAR_PBC3_lost <- dfPAR_PBR3_edit %>%
  select(-Record, -CHK, -DATAH) %>%
  filter(Date > "2024-11-11") %>%
  rename(PAR = MO) %>%
  mutate(Trial = "PBC 3", .after = Time) %>%
  mutate(Date = as.Date(Date))
  
# finish normal data wrangling for trial
dfPAR_PBC3_final <- dfPAR_PBC3_edit %>%
  select(-Record, -CHK, -DATAH) %>%
  filter(Date <= "2025/03/17") %>% #make sure dates for PBC3 reflect dates of trial
  rename(PAR = MO) %>%
  mutate(Trial = "PBC 3", .after = Time) %>%
  mutate(Date = as.Date(Date)) %>%
  # add in lost observations
  rbind(dfPAR_PBC3_lost)

#check if time has milliseconds 
dfPAR_PBC3_final %>%
  filter(str_detect(Time, "\\..*") == TRUE) %>%
  nrow() #none

#check for NA values
head(which(is.na(dfPAR_PBC3_final), arr.ind=TRUE)) #none


```

## Combine Frames
```{r data-wrangling}
# combine all dataframes above into one big dataframe by stacking the rows on top of each other - all dataframes must have the same number of columns

# combine first 3 trials which record PAR every 1 seconds
dfPAR_summer <- rbind(dfPAR_PBR1_final, dfPAR_PBC1_final, dfPAR_PBR2_final) %>%# make sure PAR values are numeric
  mutate(PAR = as.numeric(PAR)) %>%
  mutate(Time = gsub('.{3}$', ':00', Time)) %>%
  filter(PAR != 0)  %>% # PAR value of 0 with no decimal places seems to be some kind of logging error since all the other values have decimal places, and many of them are during daylight hours
  group_by(Date, Time, Trial) %>% # mean PAR per second
  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE))) 

# combine last 4 trials which record PAR every 5 seconds
dfPAR_fall_spring <- rbind(dfPAR_PBC2_final, dfPAR_PBR3_final, dfPAR_PBC3_final, dfPAR_PBR4_final) %>%
  mutate(PAR = as.numeric(PAR)) %>%
  mutate(Time = gsub('.{3}$', ':00', Time)) %>%
  filter(PAR != 0) # PAR value of 0 with no decimal places seems to be some kind of logging error since all the other values have decimal places, and many of them are during daylight hours
  
# all par values, but will keep using the separate par frames for down stream calculations
dfPAR_all <- rbind(dfPAR_summer, dfPAR_fall_spring)

# export dfPAR_all as excel file
# write_xlsx(dfPAR_all, path = "InputData/Clean/EnvironmentalData/AllPAR.xlsx")

```


# Calculations 
## Daily Light Integral (DLI)
```{r calculate-DLI}
# calculate DLI (for all the days in the file and output to a new file/table)
# calculation: 1) add all PAR readings for the day; 2) multiply by the time interval in seconds between measurements; 3) divide by 1000000 to complete the unit conversion

# create variable holding time interval between measurements in seconds, which is 1 second in summer and 5 seconds in fall/spring
time_interval_summer = 1
time_interval_fall_spring = 5

# create dataframe of the sum of all PAR measurements for each day, keeping track of trial
# summer
parSums_summer <- dfPAR_summer %>%
  group_by(Date, Trial) %>%
  summarise(daily_par_sum = sum(PAR))
# fall/spring
parSums_fall_spring <- dfPAR_fall_spring %>%
  group_by(Date, Trial) %>%
  summarise(daily_par_sum = sum(PAR))

# add a column to parSums dataframe that calculates the DLI using the formula above
#summer 
DLI_summer <- parSums_summer %>%
  mutate(DLI = daily_par_sum * time_interval_summer / 1000000)
#fall/spring
DLI_fall_spring <- parSums_fall_spring %>%
  mutate(DLI = daily_par_sum * time_interval_fall_spring / 1000000)

# DLI for all trials 
DLI_all <- rbind(DLI_summer, DLI_fall_spring)

# save DLI file
# write_xlsx(DLI_all, path = "InputData/Clean/EnvironmentalData/DLI-All.xlsx")

```

Make a bar plot (called a column plot in R due to the axes) to check that the data points are labeled under the correct trials by date. 
```{r plot-dli}
# check what DLI per trial looks like 
DLI_all %>%
  mutate(Trial = factor(Trial, levels = c("PBR 1", "PBC 1", "PBR 2", "PBC 2", "PBR 3", "PBC 3", "PBR 4"))) %>%
  ggplot(aes(x = Date, 
             y = DLI, 
             fill = Trial)) +
  geom_col() +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line = element_blank())

```


## Daily Par Values
Mean instantaneous PAR value per day 
```{r calculate daily PAR values}
# create a time tibble and use datetime column as the index; get rid of any rows with NA values in the datetime column

#create datetime column in dfPAR_all
dfPAR_all$datetime <- as.POSIXct(
  paste(dfPAR_all$Date, dfPAR_all$Time),
  format = "%Y-%m-%d %H:%M:%S",
  tz = "UTC")

lightTime <- as_tbl_time(dfPAR_all,index=datetime) %>%
  filter(is.na(datetime)==FALSE)

# calculate mean PAR value for each day, making sure we keep track of the treatments
dailyMeanPar <- lightTime %>%
  collapse_by("day") %>%
  group_by(Date, Trial) %>%
  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE))) %>%
  filter(!Trial %in% c("PBR 1", "PBC 1", "PBR 2"))

# check daily mean par over time 
dailyMeanPar %>%
  mutate(Trial = factor(Trial, levels = c("PBR 1", "PBC 1", "PBR 2", "PBC 2", "PBR 3", "PBC 3", "PBR 4"))) %>%
  ggplot(aes(x = Date, 
             y = PAR, 
             fill = Trial)) +
  geom_col()

```